---
title: "Untitled"
author: "Beatriz"
date: "February 1, 2019"
output: html_document
---

Load phyloseq and ggplot
```{r}
library("phyloseq") #; packageVersion("phyloseq")
library("ggplot2") #; packageVersion("ggplot2")
#theme_set(theme_bw())
```

1. Load data
```{r}
region <- "ITS"
sharedPathAn <- "/fs/hnas1-evs1/Dgrdi/grdi_eco/groups/bilodeaug/gir001/PIRL_working/ecoBiomics_Ian/"
biomTable <-import_biom( paste ( sharedPathAn, "OTU_tables/", region, "_otu_table.biom", sep=""))
mappingFile <- import_qiime_sample_data(paste(sharedPathAn, "OTU_tables/", region, "-mapping-file.txt", sep=""))
library(ape)
ITSmergedData <-  merge_phyloseq(biomTable, mappingFile) #if no tree
colnames(sample_data(ITSmergedData))[5] = "SampleType"
# colnames(tax_table(ITSmergedData))[8] = "Genus" 
# colnames(tax_table(ITSmergedData))[9] = "Species" 
mergedDatar  = transform_sample_counts(ITSmergedData, function(x) x / sum(x) )
mergedDatafr = filter_taxa(mergedDatar, function(x) mean(x) > 1e-3, TRUE)
####needs non zero mergedDatafrC = filter_taxa(mergedDatafr, function(x) sum(x > 3) > (0.2*length(x)), TRUE)
####needs non zero mergedDatafrf = filter_taxa(mergedDatafr, function(x) sd(x)/mean(x) > 3.0, TRUE)
mergedDatafrf = prune_taxa(taxa_sums(mergedDatafr) > 0, mergedDatafr)
```

See the data:
```{r}
sample_names(ITSmergedData)
ntaxa(ITSmergedData)
rank_names(ITSmergedData)
sample_variables(ITSmergedData)
```
```{r}
plot_bar(mergedDatafrf, "SampleType", fill="Rank3") #+ 
  #theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text(angle = 0, size=8, hjust = 0.5), plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none") 
#ggsave(paste (sharedPathAn, "/OTU_tables/COI-all-PWvsPS.pdf", sep = ""))

  plot_richness(ITSmergedData, x="ExtractionKit", color="SampleType", measures=c("Chao1", "Shannon"))

```

2. Soil samples
```{r}
ITSsoil = subset_samples(ITSmergedData, SampleType == "soil") 
ITSsoilr = transform_sample_counts(ITSsoil, function(x) x / sum(x) )
ITSsoilfr = filter_taxa(ITSsoilr, function(x) mean(x) > 1e-3, TRUE)
ITSsoilfrf = prune_taxa(taxa_sums(ITSsoilfr) > 0, ITSsoilfr)

plot_bar(ITSsoilfrf, "Rank2", fill="Rank3", facet_grid = "ExtractionKit~SampleType") + 
  #theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text(angle = 0, size=8, hjust = 0.5), plot.title = element_text(hjust = 0.5)) #+
  theme(legend.position = "none") 
#ggsave(paste (sharedPathAn, "/OTU_tables/COI-all-PWvsPS.pdf", sep = ""))
```

```{r}
ITSsoilAscor = subset_taxa(ITSsoilr, Rank2 == "p__Ascomycota")
ITSsoilAscofr = filter_taxa(ITSsoilAscor, function(x) mean(x) > 1e-3, TRUE)
ITSsoilAscofrf = prune_taxa(taxa_sums(ITSsoilAscofr) > 0, ITSsoilAscofr)

plot_bar(ITSsoilAscofrf, "Rank3", fill="Rank4", facet_grid = "ExtractionKit~SampleType") + 
  #theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text(angle = 0, size=8, hjust = 0.5), plot.title = element_text(hjust = 0.5)) #+
  theme(legend.position = "none") 
#ggsave(paste (sharedPathAn, "/OTU_tables/COI-all-PWvsPS.pdf", sep = ""))

```
  
```{r}
ITSsoilBasr= subset_taxa(ITSsoilr, Rank2 == "p__Basidiomycota")
ITSsoilBasfr = filter_taxa(ITSsoilBasr, function(x) mean(x) > 1e-3, TRUE)
ITSsoilBasfrf = prune_taxa(taxa_sums(ITSsoilBasfr) > 0, ITSsoilBasfr)

plot_bar(ITSsoilBasfrf, "Rank3", fill="Rank4", facet_grid = "ExtractionKit~SampleType") + 
  #theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text(angle = 0, size=8, hjust = 0.5), plot.title = element_text(hjust = 0.5)) #+
  theme(legend.position = "none") 
#ggsave(paste (sharedPathAn, "/OTU_tables/COI-all-PWvsPS.pdf", sep = ""))

```

```{r}
ITSsoilMucr= subset_taxa(ITSsoilr, Rank2 == "p__Mucoromycota")
ITSsoilMucfr = filter_taxa(ITSsoilMucr, function(x) mean(x) > 1e-3, TRUE)
ITSsoilMucfrf = prune_taxa(taxa_sums(ITSsoilMucfr) > 0, ITSsoilMucfr)

plot_bar(ITSsoilMucfrf, "Rank3", fill="Rank4", facet_grid = "ExtractionKit~SampleType") + 
  #theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text(angle = 0, size=8, hjust = 0.5), plot.title = element_text(hjust = 0.5)) #+
  theme(legend.position = "none") 
#ggsave(paste (sharedPathAn, "/OTU_tables/COI-all-PWvsPS.pdf", sep = ""))

```

```{r}
#plot_richness(ITSsoil, x="ExtractionKit", measures=c("Chao1", "Shannon"))
plot_richness(ITSsoil, x="ExtractionKit", color="SampleType", measures=c("Chao1", "Shannon"))
```
Even though we see that diversity of PowerWater seems higher, we dont have a p-value to know if PW is significantly higher than FD


3. Benthic samples
```{r}
ITSben = subset_samples(ITSmergedData, SampleType == "benthic") 
ITSbenr = transform_sample_counts(ITSben, function(x) x / sum(x) )
ITSbenfr = filter_taxa(ITSbenr, function(x) mean(x) > 1e-3, TRUE)
ITSbenfrf = prune_taxa(taxa_sums(ITSbenfr) > 0, ITSbenfr)

plot_bar(ITSbenfrf, "Rank4", fill="Rank5", facet_grid = "ExtractionKit~SampleType") + 
  #theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text(angle = 0, size=8, hjust = 0.5), plot.title = element_text(hjust = 0.5)) #+
  theme(legend.position = "none") 
#ggsave(paste (sharedPathAn, "/OTU_tables/COI-all-PWvsPS.pdf", sep = ""))
```

```{r}
plot_richness(ITSben, x="ExtractionKit", color="SampleType", measures=c("Chao1", "Shannon"))
```
Even though we see that diversity of PowerWater seems higher, we dont have a p-value to know if PW is significantly higher than FD

------------------------------------------COMPARE ITS APPROACHES--------------------------
The aim is to compare de novo using rdp or blast classification, and closed reference (only rdp available). First lets compare diversity, then compare taxa graphically. To ease the visualizations we will: 
1. Look at benthic samples only
2. Remove everything that is unclassified from genus onwards. 


```{r}
mergedDatafrf2<- subset_taxa(mergedDatafrf, Rank5!="NA" )
mergedDatafrf2<- subset_taxa(mergedDatafrf2, Rank5!="f__unidentified" )
# UmergedDatafrF <- subset_taxa(mergedDatafrf, Rank5=="NA" | Rank5== "f__unidentified" )
# plot_bar(UmergedDatafrF, "Rank3", facet_grid = "SampleType", fill="Rank7") #+ 
# plot_bar(UmergedDatafrF, "Rank5", facet_grid = "SampleType", fill="Rank6")
# UmergedDatafrF = prune_taxa(taxa_sums(UmergedDatafrF) > 0, UmergedDatafrF)
ITSben2 = subset_samples(mergedDatafrf2, SampleType == "benthic") 
```

```{r}
region <- "ITS"
sharedPathAn <- "/fs/hnas1-evs1/Dgrdi/grdi_eco/groups/bilodeaug/gir001/PIRL_working/ecoBiomics_Ian/"
biomTableCR <-import_biom( paste ( sharedPathAn, "other-otu-methods/closed-ref/", region, "_otu_table.biom", sep=""))
mappingFile <- import_qiime_sample_data(paste(sharedPathAn, "OTU_tables/", region, "-mapping-file.txt", sep=""))
ITSmergedDataCR <-  merge_phyloseq(biomTableCR, mappingFile) #if no tree
colnames(sample_data(ITSmergedDataCR))[5] = "SampleType"
mergedDatarCR  = transform_sample_counts(ITSmergedDataCR, function(x) x / sum(x) )
mergedDatafrCR = filter_taxa(mergedDatarCR, function(x) mean(x) > 1e-3, TRUE)
mergedDatafrfCR = prune_taxa(taxa_sums(mergedDatafrCR) > 0, mergedDatafrCR)
mergedDatafrfCR2<- subset_taxa(mergedDatafrfCR, Rank5!="NA" )
mergedDatafrfCR2<- subset_taxa(mergedDatafrfCR2, Rank5!="f__unidentified" )
ITSbenCR2 = subset_samples(mergedDatafrfCR2, SampleType == "benthic") 
```

```{r}
region <- "ITS"
sharedPathAn <- "/fs/hnas1-evs1/Dgrdi/grdi_eco/groups/bilodeaug/gir001/PIRL_working/ecoBiomics_Ian/"
biomTableB <-import_biom( paste ( sharedPathAn, "K_classify/ITS-blast/", region, "_otu_table.biom", sep=""))
mappingFile <- import_qiime_sample_data(paste(sharedPathAn, "OTU_tables/", region, "-mapping-file.txt", sep=""))
ITSmergedDataB <-  merge_phyloseq(biomTableB, mappingFile) #if no tree
colnames(sample_data(ITSmergedDataB))[5] = "SampleType"
mergedDatarB  = transform_sample_counts(ITSmergedDataB, function(x) x / sum(x) )
mergedDatafrB = filter_taxa(mergedDatarB, function(x) mean(x) > 1e-3, TRUE)
mergedDatafrfB = prune_taxa(taxa_sums(mergedDatafrB) > 0, mergedDatafrB)
mergedDatafrfB2<- subset_taxa(mergedDatafrfB, Rank5!="NA" )
mergedDatafrfB2<- subset_taxa(mergedDatafrfB2, Rank5!="f__unidentified" )
ITSbenB2 = subset_samples(mergedDatafrfB2, SampleType == "benthic") 

#NAmergedDatafrfB2<- subset_taxa(mergedDatafrfB, Rank5=="NA" |  Rank5=="f__unidentified" )
```

For plots:

```{r}
library(grid)
p1 <- plot_richness(ITSmergedData, x="SampleType", color="ExtractionKit", measures=c("Observed", "Chao1")) + 
  ggtitle("De novo OTU Picking using the RDP classifier") + theme(plot.title = element_text(hjust = 0.5, size = 10), 
                                                                  axis.title=element_text(size=8)) 
p2 <- plot_richness(ITSmergedDataB, x="SampleType", color="ExtractionKit", measures=c("Observed", "Chao1")) + 
  ggtitle("De novo OTU Picking using BLAST") + theme(plot.title = element_text(hjust = 0.5, size = 10),
                                                     axis.title=element_text(size=8))
p3 <- plot_richness(ITSmergedDataCR, x="SampleType", color="ExtractionKit", measures=c("Observed", "Chao1")) + 
  ggtitle("Closed reference OTU Picking using the RDP classifier") + theme(plot.title = element_text(hjust = 0.5, size = 10),
                                                                           axis.title=element_text(size=8))
grid.newpage()
grid.draw(rbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(p3), size = "last"))

plot_bar(ITSben2, "Rank5",  fill="Rank7") + 
  theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text( size=8 ), plot.title = element_text(hjust = 0.5)) +
  ggtitle("De novo OTU Picking using the RDP classifier")  #+ theme(legend.position = "none") 

plot_bar(ITSbenCR2, "Rank5",  fill="Rank7") + 
  ggtitle("Closed reference OTU Picking using the RDP classifier")  + theme(legend.position = "none") 

plot_bar(ITSbenB2, "Rank5", fill="Rank7") + 
  theme(legend.key.size = unit(0.3,"cm"), axis.text.x = element_text( size=8 ), plot.title = element_text(hjust = 0.5)) +
  ggtitle("De novo OTU Picking using BLAST") #+ theme(legend.position = "none") 

  

```

