---
title: "Maker_Pipeline"
author: "Emily Giroux"
date: "12/6/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Getting started in R:** 
Set the working directory > setwd("~/") Check version installed
```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy = TRUE, fig.align='center')
```

**This will help us when finding our files to source functions:**
```{r sourcing_my_functions}
install.packages("rprojroot")
library(rprojroot)
# We specify ours is an RStudio project
# The root object contains a function that will help us locate our package r files regardless of our current working directory
root <- rprojroot::is_rstudio_project
scriptsPath <- root$make_fix_file(".")("R")
scripts  <- dir(root$find_file("R", path = root$find_file()))
scriptsl <- paste(scriptsPath, scripts, sep = "//")
lapply(scriptsl, source)
```

**Packages to install and have ready for later analyses:**
```{r}
install.packages("data.table")
```

**User: Define the path to the shared folder where the main working directory will be.**
```{r setting_the_main_directory, cache=TRUE}
sharedPath <- "/isilon/cfia-ottawa-fallowfield/users/girouxeml/PIRL_working_directory/"
```

**Define where the analysis path is, or the path to the current project:**
```{r setting_the_main_directory, cache=TRUE}
analysis   <- "Fusarium_Pramorum_species_GenomeAn_IonTorrent_2018/"
sharedPathAn <- paste(sharedPath, analysis, sep="")
```

**Define path variables to programs and scripts used:**
```{r}
# Biocluster system-wide programs:
augustBioCPath  <- "/isilon/biodiversity/pipelines/maker-2.10/augustus.2.7/bin/augustus"
augEvalBioCPath <- "/isilon/biodiversity/pipelines/maker-2.10/augustus.2.7/bin/augustus"
bedtoolsPath    <- "/opt/bio/BEDTools/bin/bedtools"
blastallPath    <- "/opt/bio/ncbi/bin/blastall"
etrainBioCPath  <- "/isilon/biodiversity/pipelines/maker-2.10/augustus.2.7/bin/etraining"
optimAugustPath <- "/isilon/biodiversity/pipelines/maker-2.10/augustus.2.7/scripts/optimize_augustus.pl"
pathtRNA_scan   <- "/opt/bio/tRNAscan-SE/bin/tRNAscan-SE"
# randomSplitPath <- "/isilon/biodiversity/pipelines/maker-2.10/augustus.2.7/scripts/randomSplit.pl"
# newSpeciesPath  <- "/isilon/biodiversity/pipelines/maker-2.10/augustus.2.7/scripts/new_species.pl"


# *** Revisit this organization. Perhaps best that all programs be /home/ and databases be in local cfia-ottawa?
# CFIA-ACIA users home directory programs:
progPath        <- "/home/CFIA-ACIA/girouxeml/prog/"
aedCDFgenePath  <- paste(progPath, "scripts_pl/AED_cdf_generator.pl", sep = "")
blastpPath      <- paste(progPath, "miniconda/bin/blastp", sep = "")
buscoPath       <- paste(progPath, "busco/scripts/run_BUSCO.py", sep = "")
fathomPath      <- paste(progPath, "snap/fathom", sep = "")
forgePath       <- paste(progPath, "snap/forge", sep = "")
gagPath         <- paste(progPath, "gag/genomeannotation-GAG-40ea515/gag.py", sep = "")
genemarkPath    <- paste(progPath, "genemark-es/gmes_petap.pl", sep = "") 
hmmAssemPath    <- paste(progPath, "snap/hmm-assembler.pl", sep = "") 
iprUpdateMaker  <- paste(progPath, "maker/bin/ipr_update_gff", sep = "")
iprscan2gff3    <- paste(progPath, "maker/bin/iprscan2gff3", sep = "")
jbrowseFlatfile2json <- paste(progPath, "jbrowse/JBrowse-1.12.3/bin/flatfile-to-json.pl", sep = "")
jbrowsePrepRefsSeqsPath <- paste(progPath, "jbrowse/JBrowse-1.12.3/bin/prepare-refseqs.pl", sep = "") 
makerDatMapPath <- paste(progPath, "maker/bin/map_data_ids", sep = "")
makerMapFasta   <- paste(progPath, "maker/bin/map_fasta_ids", sep = "")
makerFastamerge <- paste(progPath, "maker/bin/fasta_merge", sep = "") 
makerFuncFasta  <- paste(progPath, "maker/bin/maker_functional_fasta", sep = "") 
makerFuncGff    <- paste(progPath, "maker/bin/maker_functional_gff", sep = "")
makerGFF3merge  <- paste(progPath, "maker/bin/gff3_merge", sep = "") 
makerIPR2gff3   <- paste(progPath, "maker/bin/iprscan2gff3", sep = "")
makerIPRupdate  <- paste(progPath, "maker/bin/ipr_update_gff", sep = "")
makerMapGffPath <- paste(progPath, "maker/bin/map_gff_ids", sep = "")
makerMapPath    <- paste(progPath, "maker/bin/maker_map_ids", sep = "")
makerPath       <- paste(progPath, "maker/bin/maker", sep = "") 
maker2zffPath   <- paste(progPath, "maker/bin/maker2zff", sep = "") 
pathtRNA_scan   <- paste(progPath, "tRNAscan-SE/bin/tRNAscan-SE", sep = "") 
processRepeatsPath <- paste(progPath, "RepeatMasker/ProcessRepeats", sep = "") 
repMaskerPath   <- paste(progPath, "RepeatMasker/RepeatMasker", sep = "") 
repModBuildDBPath  <- paste(progPath, "RepeatModeler-open-1.0.11/BuildDatabase", sep = "") 
repModlerPath   <- paste(progPath, "RepeatModeler-open-1.0.11/RepeatModeler", sep = "") 
rmOutToGFF3Path <- paste(progPath, "RepeatMasker/util/rmOutToGFF3.pl", sep = "")
tbl2asnPath     <- paste(progPath, "linux64.tbl2asn", sep = "")
scriptsPath     <- paste(progPath, "scripts_pl/", sep = "")
fixGAGNamePath  <- paste(scriptsPath, "fix_GAG_Name.sh", sep = "")
zff2augGbkPath   <- paste(scriptsPath, "zff2augustus_gbk.pl", sep = "")

# CFIA-Ottawa-Fallowfield user directory programs:
programsPath    <- "/isilon/cfia-ottawa-fallowfield/users/girouxeml/prog/"
buscoPezDataSet <- paste(programsPath, "busco_datasets/pezizomycotina_odb9/", sep = "")
buscoSorDataSet <- paste(programsPath, "busco_datasets/sordariomyceta_odb9", sep = "")
buscoFunDataSet <- paste(programsPath, "busco_datasets/fungi_odb9", sep = "")
interproPath    <- paste(programsPath, "my_interproscan/interproscan-5.24-63.0/interproscan.sh", sep = "")
uniProtSwissPDB <- paste(programsPath, "databases/uniprot_sprot.fasta", sep = "")
```

**Define paths to reference files used:**  
**Note,** require entire proteome from a  min of two related species and perhaps all of UniProt/SwissProt. 
See advice for multiple proteomes for homology: https://groups.google.com/forum/#!topic/maker-devel/jbBm_4ycFU8
  
**_Marsonnina brunnea_** available at: https://www.ncbi.nlm.nih.gov/genome/?term=txid698440[orgn]:
ftp://ftp.ncbi.nlm.nih.gov/genomes/all/GCA/000/298/775/GCA_000298775.1_ASM29877v1  
  
For **_Sclerotinia sclerotiorum_**: https://www.ncbi.nlm.nih.gov/genome/?term=txid5180[orgn]   
  
All others available from ensemble. Retrieved them based on filtering:  
*http://www.uniprot.org/proteomes/?query=reference%3Ayes+AND+taxonomy%3A%22Eukaryota+%5B2759%5D%22+AND+taxonomy%3AHelotiales&sort=score*
```{r}
referencesPath  <- paste(sharedPath, "References/", sep = "")

cdnaFuspoaPath  <- paste(referencesPath, "Fusarium_poae_FPOA1.0/Fusarium_poae.FPOA1.0.cdna.all.fa", sep = "")
pepFuspoaPath   <- paste(referencesPath, "Fusarium_poae_FPOA1.0/Fusarium_poae.FPOA1.0.pep.all.fa", sep = "")
cdsFuspoaPath   <- paste(referencesPath, "Fusarium_poae_FPOA1.0/Fusarium_poae.FPOA1.0.cds.all.fa", sep = "")
dnaFuspoaPath   <- paste(referencesPath, "Fusarium_poae_FPOA1.0/Fusarium_poae.FPOA1.0.dna.toplevel.fa", sep = "")
gff3FuspoaPath  <- paste(referencesPath, "Fusarium_poae_FPOA1.0/Fusarium_poae.FPOA1.0.40.gff3", sep = "")

pepFuspoa2Path   <- paste(referencesPath, "Fuspoae_GCA_001675295.1/Fusarium_poae_GCA_001675295.1_FPOA1.0_protein.faa", sep = "")
dnaFuspoa2Path   <- paste(referencesPath, "Fuspoae_GCA_001675295.1/Fusarium_poae_GCA_001675295.1_FPOA1.0_genomic.fna", sep = "")
gff3Fuspoa2Path  <- paste(referencesPath, "Fuspoae_GCA_001675295.1/Fusarium_poae_GCA_001675295.1_FPOA1.0_genomic.gff", sep = "")

dnaFusavePath   <- paste(referencesPath, "Fusave_GCA_000769215.1/Fusarium_avenaceum_GCA_000769295.1_Fave_LH27_genomic.fna", sep = "")
pepFusavePath   <- paste(referencesPath, "Fusave_GCA_000769215.1/Fusarium_avenaceum_GCA_000769215.1_Fave_05001_protein.faa", sep = "")
gffFusavePath   <- paste(referencesPath, "Fusave_GCA_000769215.1/Fusarium_avenaceum_GCA_000769215.1_Fave_05001_genomic.gff", sep = "")

dnaFuslangPath  <- paste(referencesPath, "Fuslang_GCA_001292635.1/Fusarium_langsethiae_GCA_001292635.1_ASM129263v1_genomic.fna", sep = "")
pepFuslangPath  <- paste(referencesPath, "Fuslang_GCA_001292635.1/Fusarium_langsethiae_GCA_001292635.1_ASM129263v1_protein.faa", sep = "")
gffFuslangPath  <- paste(referencesPath, "Fuslang_GCA_001292635.1/Fusarium_langsethiae_GCA_001292635.1_ASM129263v1_genomic.gff", sep = "")

dnaFusmangPath <- paste(referencesPath,"Fusmang_GCA_900044065.1/Fusarium_mangiferae_GCA_900044065.1_Genome_assembly_version_1_genomic.fna", sep = "")
pepFusmangPath <- paste(referencesPath,"Fusmang_GCA_900044065.1/Fusarium_mangiferae_GCA_900044065.1_Genome_assembly_version_1_protein.faa", sep = "")
gffFusmangPath <- paste(referencesPath,"Fusmang_GCA_900044065.1/Fusarium_mangiferae_GCA_900044065.1_Genome_assembly_version_1_genomic.gff", sep = "")

dnaFusfujPath   <- paste(referencesPath,"Fusfuj_GCA_900096685.1/Fusarium_fujikuroi_GCA_900096685.1_version_1_genomic.fna", sep = "")
pepFusfujPath   <- paste(referencesPath,"Fusfuj_GCA_900096685.1/Fusarium_fujikuroi_GCA_900096685.1_version_1_protein.faa", sep = "")
gffFusfujPath   <- paste(referencesPath,"Fusfuj_GCA_900096685.1/Fusarium_fujikuroi_GCA_900096685.1_version_1_genomic.gff", sep = "")

dnaFusgramPath   <- paste(referencesPath,"Fusgram_GCA_002352725.1/Fusarium_graminearum_GCA_002352725.1_FGRA07_genomic.fna", sep = "")
pepFusgramPath   <- paste(referencesPath,"Fusgram_GCA_002352725.1/Fusarium_graminearum_GCA_002352725.1_FGRA07_protein.faa", sep = "")
gffFusgramPath   <- paste(referencesPath,"Fusgram_GCA_002352725.1/Fusarium_graminearum_GCA_002352725.1_FGRA07_genomic.gff", sep = "")

dnaFusoxysPath   <- paste(referencesPath,"Fusoxys_GCA_000271705.2/Fusarium_oxysporum_GCA_000271705.2_FO_Fo47_V1_genomic.fna", sep = "")
pepFusoxysPath   <- paste(referencesPath,"Fusoxys_GCA_000271705.2/Fusarium_oxysporum_GCA_000271705.2_FO_Fo47_V1_protein.faa", sep = "")
gffFusoxysPath   <- paste(referencesPath,"Fusoxys_GCA_000271705.2/Fusarium_oxysporum_GCA_000271705.2_FO_Fo47_V1_genomic.gff", sep = "")

dnaFusprolPath   <- paste(referencesPath,"Fusprol_GCA_003290285.1/Fusarium_proliferatum_GCA_003290285.1_FPRO05_genomic.fna", sep = "")
pepFusprolPath   <- paste(referencesPath,"Fusprol_GCA_003290285.1/Fusarium_proliferatum_GCA_003290285.1_FPRO05_protein.faa", sep = "")
gffFusprolPath   <- paste(referencesPath,"Fusprol_GCA_003290285.1/Fusarium_proliferatum_GCA_003290285.1_FPRO05_genomic.gff", sep = "")

dnaFusvertPath   <- paste(referencesPath,"Fusvert_GCA_003317015.1/Fusarium_verticilliodes_GCA_003317015.1_ASM331701v1_genomic.fna", sep = "")
pepFusvertPath   <- paste(referencesPath,"Fusvert_GCA_003317015.1/Fusarium_verticilliodes_GCA_003317015.1_ASM331701v1_protein.faa", sep = "")
gffFusvertPath   <- paste(referencesPath,"Fusvert_GCA_003317015.1/Fusarium_verticilliodes_GCA_003317015.1_ASM331701v1_genomic.gff", sep = "")

dnaFussolPath   <- paste(referencesPath,"Fussol_GCA_000151355.1/Fusarium_solani_GCA_000151355.1_v2.0_genomic.fna", sep = "")
pepFussolPath   <- paste(referencesPath,"Fussol_GCA_000151355.1/Fusarium_solani_GCA_000151355.1_v2.0_protein.faa", sep = "")
gffFussolPath   <- paste(referencesPath,"Fussol_GCA_000151355.1/Fusarium_solani_GCA_000151355.1_v2.0_genomic.gff", sep = "")

dnaFuspseuPath   <- paste(referencesPath,"Fuspseu_GCA_000303195.2/Fusarium_pseudograminearum_GCA_000303195.2_FP7_genomic.fna", sep = "")
pepFuspseuPath   <- paste(referencesPath,"Fuspseu_GCA_000303195.2/Fusarium_pseudograminearum_GCA_000303195.2_FP7_protein.faa", sep = "")
gffFuspseuPath   <- paste(referencesPath,"Fuspseu_GCA_000303195.2/Fusarium_pseudograminearum_GCA_000303195.2_FP7_genomic.gff", sep = "")
```

**Read in the metadata table:**  
We need to specify where we put the assembly files for our genomes.
We can either read in the table, or take what we have from the environment. Reading in the table is better.
```{r}
library(data.table)
metadataAssemblies <- fread(paste(referencesPath, "fustric_metadata_Assembly_GCA_900382705.2sep2018.csv", sep = ""),
                            sep = "auto", header = TRUE)
metadataAssemblies[, V1:=NULL]
```

### 1. De novo Repeat Identification.  
The first, and very important, step to genome annotation is identifying repetitive content. Existing libraries 
from Repbase or from internal efforts are great, but it is also important to identify repeats de novo from your reference 
genome using RepeatModeler. This is pretty easy to do and normally only takes a couple days using 8-12 cores.  
Why do this? 
If we were working with a well-studied organism, we wouldn't need to, we could just stick to the database that repeatmasker 
used which we updated with RepBase. But our species are not well-studied *Lachnellula*, and so we need to create 
a species-specific repeat library for each of our species.  
```{r}
library(dplyr)
metadataAssemblies <- filter(metadataAssemblies, SppAbbrv == "Fustric")
#setnames(metadataAssemblies, "SppAbbr", "SppAbbrv")
```
#### 1-A. Setting up directories.
**Make a directory within each species-specific directory that will contain all Repeat and MAKER annotation data:**
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(sharedPathAn, metadataAssemblies$ScientificName[k], "/", 
                     paste(metadataAssemblies$SppAbbrv[k], "RepeatDB", sep = "_"),
                     sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$RepeatDBPath <- paste(sharedPathAn, metadataAssemblies$ScientificName, "/",
                                         metadataAssemblies$SppAbbrv, "_RepeatDB/", sep = "")
```

#### 1-B. Build a repeat database for RepeatModeler for each species.
```{r}
prefix <- "buildDB_fustric"
cmd <- with(metadataAssemblies, 
            paste("cd ", RepeatDBPath, " && ", repModBuildDBPath, " -name ", SppAbbrv, " ", FixedNewAssemName, sep = ""))
cmd
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```
To remove the output files after you are done:
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Run RepeatModeler using the databases just made for each species:**
```{r}
node   <- 6
prefix <- "repeatModeler_Fustric"

cmd <- with(metadataAssemblies,
            paste("cd ", RepeatDBPath, " && ", repModlerPath, " -pa ", node, " -database ", 
                  SppAbbrv, " 2>&1 | tee repeatmodeler.log ", sep = ""))
cmd
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```
qsub -pe smp 20 -cwd -S /bin/bash "name of bash file"
**Note:**
Remember to use the modified assemblies that have the extra details that Newbler added after the >contig 
name (3 tabs for contig name, length and some other detail) removed because this info created problems downstream.    
  
**Note:** if there is an unexpected system failure, RepeatModeler can recover from it and continue from where it
previously left off. See the README file in its program directory.  

The next step is to copy all the consensi.fa.classified libraries to a directory and combine them, then separate 
fasta entries with known and unknown repeats. 
  
**Setting things up:**
```{r}
speciesRepeatDBname <- "spp_Fus_repeatDB"
dir.create(paste(sharedPathAn, "RepeatDatabases", sep = ""), showWarnings = TRUE, recursive = FALSE)
projRepeatDB <- paste(sharedPathAn, "RepeatDatabases/", sep = "")

dir.create(paste(projRepeatDB, speciesRepeatDBname, sep = ""), showWarnings = TRUE, recursive = FALSE)
allSppRepeatDBpath <- paste(projRepeatDB, speciesRepeatDBname, "/", sep = "")

for(i in 1:nrow(metadataAssemblies)){
    tmp[i] <- list.files(path = metadataAssemblies$RepeatDBPath[i], pattern = "^RM_.*")
    metadataAssemblies$RepeatModelerDir[i] <- tmp[i]
    metadataAssemblies$RepeatModelerPath[i] <- paste(metadataAssemblies$RepeatDBPath[i], 
                                                     metadataAssemblies$RepeatModelerDir[i], "/", sep = "")}
```

**Copying the files:**
```{r}
prefix <- "copy_Spp_RepModeler"
cmd <- with(metadataAssemblies, 
            paste("cp ", paste(metadataAssemblies$RepeatModelerPath, "consensi.fa.classified", sep = ""),
                  " ", paste(sharedPathAn, "RepeatDatabases/spp_Fus_repeatDB/", 
                             metadataAssemblies$SppAbbrv, "_consensi.fa.classified", sep = "")))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**Combine them into one file:**
```{r}
metadataAssemblies$repModelerConsensi <- paste(metadataAssemblies$SppAbbrv, "_consensi.fa.classified", sep = "")
list   <- metadataAssemblies$repModelerConsensi
output <- paste(list, collapse = " ")
# I'm automating the output file name, since all the species in this set (*Fusarium spp*), are within the same
# genus, but if this is not the case for another set, specify the output file name manually.
species <- c(metadataAssemblies$ScientificName)
genus   <- gsub("_.*", "", species)
genus   <- unique(genus)
repModAllsppFileName <- paste(genus, "consensi.fa.classified", sep = "_")
cmd <- paste("cat ", output, " > ", genus, "_", "consensi.fa.classified", sep = "")
write(cmd, file = paste(allSppRepeatDBpath, repModAllsppFileName, sep = ""))
```

**Split unknown and known repeats to separate files:**     
We'll use perl scripts for this (in ~/prog/scripts_pl/): _selectFasta_byIDUnknown.pl_, _selectFasta_byID_Known.pl_    
  
**Note:** You need to open these perl scripts and edit them to fix the names you want for input and output,
and you need to run them in the directory where your input is.
```{r}
allSppRepeatDBpath <- paste(projRepeatDB, speciesRepeatDBname, "/", sep = "")
# Write script to parse and get all repeats that are "Known"
identity <- 1
infile   <- repModAllsppFileName
parseFastaByHeaderPattern(infile, identity)

# Repeat the above with identity <- 2, to get all the unknown fastas:
identity <- 2
infile   <- repModAllsppFileName
parseFastaByHeaderPattern(infile, identity)
```

#### 1-C. Quality check of the repeat database built with RepeatModeler.
QC the output from the Unknown_Fusarium_consensi.fa.classified repeat library to check for repeats that 
are really genes (gene families) or RNA features:

**Blastx vs nr:**  
I did this one on the command line. Blastx vs nr for protein coding genes (I qlogin'd 
with 22 processors, so thread using 21): 
  
**_$ blastx -db /isilon/ottawa-rdc/reference/ncbi/blastdb/reference/nr/nr -query unknown_Fusarium_consensi.fa.classified -evalue 10e-5 -num_alignments 5 -num_threads 21 > Unknown_Fusarium_consensi.fa.classified.nrcheck.blast.out_**  

To look at the output, search for "Sequences producing significant alignments". For all those found, remove those sequences from the  
unknown_Fusarium_consensi.fa.classified file, and rename and save that file as:   
unknown_Fusarium_consensi.fa.nr.checked.blast.classified.   

**Blastn vs RNA database for ribosomal or other RNA genes:**  
Blastn vs RNA database for ribosomal or other RNA genes. Rfam.fasta comes with the Rfam download and it requires you to
download the Rfam database. Do this only when it hasn't already been done, or if it's been a really
long time and you need to update the database:   
  
**_$ cd ~/prog/_**  
**_$ mkdir rfam_**  
**_$ cd rfam_**  
**_$ wget -r_ ftp://ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/fasta_files/**  
**_$ mv -f_ ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/fasta_files/\* .**  
**_$ gunzip RF0\*_**  
**_$ cat *.fa > rfam.fasta_**  
**_$ mv RF\*_ ftp.ebi.ac.uk/pub/databases/Rfam/CURRENT/fasta_files/**  
**_$ makeblastdb -in rfam.fasta -dbtype nucl_**    

**Run Blastn against the rfam database in the Fusarium_species/RepeatDatabases/spp_Fus_repeatDB dir:**
```{r}
prefix <- "RepeatDB_blastn_Fustric"
cmd    <- paste("blastn -db /home/CFIA-ACIA/girouxeml/prog/rfam/rfam.fasta -query ",
                paste(allSppRepeatDBpath, "Unknown_", infile, sep = ""),
                " -evalue 10e-5 -num_alignments 5 -num_threads 20 > ",
                paste(allSppRepeatDBpath, "Unknown_", infile, ".rnacheck.blast.out", sep = ""),
                sep = "")
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Look at the output from the Blastx and Blastn runs, and remove repeats that are actually really genes (gene families) or RNA features.
I did not find any hits matching in the Rfam run, so nothing to remove from there. I did find blast hits with the nr check. I removed 
those repeats with good hits. 
The final unknown file is called "Unknown_Fusarium_consensi.fa.nr.rna.checked.blast.classified" and the known repeat library file
is called "Known_Fusarium_consensi.fa.classified".

**Convert the EMBL repeat library from RepeatMasker to fasta format:**  
Place the repeat library in its own dir in the Fusarium_species dir:
```{r}
dir.create(paste(projRepeatDB, "embl_RepeatMaskerDB_Fustric", sep = ""), showWarnings = TRUE, recursive = FALSE)

prefix <- "embl_buildRepMaskeLib_Fustric"
cmd    <- paste("perl /home/CFIA-ACIA/girouxeml/prog/RepeatMasker/util/buildRMLibFromEMBL.pl ",
                "/home/CFIA-ACIA/girouxeml/prog/RepeatMasker/Libraries/RepeatMaskerLib.embl ",
                paste(allSppRepeatDBpath, "known_Fusarium_consensi.fa.classified", sep = ""),
                " ", paste(allSppRepeatDBpath, "unknown_Fusarium_consensi.fa.nr.rna.checked.blast.classified", sep = ""),
                " > ", paste(projRepeatDB, "embl_RepeatMaskerDB/RepeatMaskerLib.fasta", sep = ""),
                sep = "")
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Further steps can be taken to annotate the resulting library, but the most important reason for this 
library is for downstream gene prediction. This step came from an example where a Boa library was 
combined with several other snakes and annotated:  
  
https://gist.github.com/darencard/bb1001ac1532dd4225b030cf0cd61ce2  
  
#### 1-D. Full Repeat Annotation with RepeatMasker:   
  
**Full Repeat Annotation:**  
Depending on the species, the de novo library can be fed right into MAKER, however, based on the very thorough example provided 
by https://gist.github.com/darencard/bb1001ac1532dd4225b030cf0cd61ce2, we'll be doing a more complex repeat identification.
  
**EMBL repeat library masking:**    
First, we mask using a currated library to overcome a previously-identified issue with the Repbase annotation. 
I used the known library from repeatmasker.lib.EMBL, which I converted to fasta format using RepeatMasker/util/buildRMLibFromEMBL.pl script.
Then we pass the species genome assemblies through the RepeatMaskerLib.fasta using RepeatMasker.
```{r}
for(i in 1:length(metadataAssemblies$RepeatDBPath)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[i], "embl_mask_Fustric", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$embl_mask_path <- paste(metadataAssemblies$RepeatDBPath, "embl_mask_Fustric/", sep = "")

prefix <- "embl_mask_Fustric"
node   <- 12
cmd    <- with(metadataAssemblies,
               paste("cd ", embl_mask_path, " && ", repMaskerPath, " -pa ", node, " -lib ",
                     paste(projRepeatDB, "embl_RepeatMaskerDB/RepeatMaskerLib_Fustric.fasta ", sep = ""),
                     " -dir . ", FixedNewAssemName, sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Known repeat library masking:**    
Then the masked FASTA from this search can be used as input for the next search. I also normally 
rename the outputs after each round so they are more representative of what they contain. Do 2 rounds, 
using the known and unknown *Fusarium* repeat libraries created, split, so that the known elements 
would be preferentially annotated over unknown, to the degree possible.  
```{r}
for(i in 1:length(metadataAssemblies$RepeatDBPath)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[i], "Known_Mask_Fustric", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$Known_Mask_path <- paste(metadataAssemblies$RepeatDBPath, "Known_Mask_Fustric/", sep = "")

prefix <- "Known_Mask_Fustric"
node   <- 12
cmd    <- with(metadataAssemblies,
               paste("cd ", Known_Mask_path, " && ", repMaskerPath, " -pa ", node, " -lib ",
                     paste(allSppRepeatDBpath, "known_Fusarium_consensi.fa.classified", sep = ""),
                     " -dir . ", paste(embl_mask_path, "*.fna.masked", sep = ""), sep = ""))
# Run RepeatMasker with known repeat lib 1st, using as input the embl masked output
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Unknown repeat library masking:**    
Run RepeatMasker with unknown repeat lib, using as input the unknown masked output. Note, make sure to use the unknown lib
that was passed through NR, RNA, and BLAST checks to remove repeats that were actually part of genes or functional RNAs.
```{r}
for(i in 1:length(metadataAssemblies$RepeatDBPath)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[i], "Unknown_Mask_Fustric", sep = ""), showWarnings = TRUE, recursive = FALSE)
}
metadataAssemblies$Unknown_Mask_path <- paste(metadataAssemblies$RepeatDBPath, "Unknown_Mask_Fustric/", sep = "")

# Run RepeatMasker with unknown repeat lib 2nd, using as input the Known masked output
prefix <- "Unknown_Mask_Fustric"
node   <- 12
cmd    <- with(metadataAssemblies,
               paste("cd ", Unknown_Mask_path, " && ", repMaskerPath, " -pa ", node, " -lib ",
                     paste(allSppRepeatDBpath, "unknown_Fusarium_consensi.fa.nr.rna.checked.blast.classified", sep = ""),
                     " -dir . ", paste(Known_Mask_path, "*.fna.masked.masked", sep = ""), sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
# Ignore the segmentation fault errors, these are tries over already masked strings
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Finally, results from each round must be analyzed together to produce the final repeat annotation:**
```{r}
# Make a directory in each species dir for the repeats masked with the embl lib:
for(i in 1:length(metadataAssemblies$RepeatDBPath)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[i], "Full_mask_Fustric", sep = ""), showWarnings = TRUE, recursive = FALSE)}

prefix <- "copy_masked_rep_fna_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cp ", paste(RepeatDBPath, "Unknown_Mask_Fustric/*.fna.masked.masked.masked ", sep = ""),
                     paste(RepeatDBPath, "Full_mask_Fustric/", SppAbbrv, ".assembly.full_mask.fa", sep = ""),
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

prefix2 <- "copy_masked_rep_out_Fustric"
cmd2    <- with(metadataAssemblies,
                paste("cp ", paste(RepeatDBPath, "Unknown_Mask_Fustric/*.fna.masked.masked.out ", sep = ""),
                      paste(RepeatDBPath, "Full_mask_Fustric/", SppAbbrv, ".assembly.full_mask.out", sep = ""),
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2)
```

**Gunzip and then cat repeat outputs:**
```{r}
prefix <- "gunzipToCatRepeatFasta_Fustric"
cmd    <- with(metadataAssemblies, 
               paste("gunzip ", 
                     paste(RepeatDBPath, "embl_mask_Fustric/*.cat.gz ", sep = ""),
                     paste(RepeatDBPath, "Known_Mask_Fustric/*.cat.gz ", sep = ""),
                     paste(RepeatDBPath, "Unknown_Mask_Fustric/*.cat.gz ", sep = ""),
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Make a directory in each species dir for the final repeats masked:**
```{r}
for(i in 1:length(metadataAssemblies$RepeatDBPath)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[i], "Final_mask_Fustric", sep = ""), showWarnings = TRUE, recursive = FALSE)
}

prefix <- "catRepeatFasta_Fustric"
cmd    <- with(metadataAssemblies, 
               paste("cat ", 
                     paste(RepeatDBPath, "embl_mask_Fustric/*.cat ", sep = ""),
                     paste(RepeatDBPath, "Known_Mask_Fustric/*.cat ", sep = ""),
                     paste(RepeatDBPath, "Unknown_Mask_Fustric/*.cat ", sep = ""),
                     " > ", paste(RepeatDBPath, "Final_mask_Fustric/", SppAbbrv, ".assembly.full_mask.cat", sep = ""),
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Process the repeats:**
```{r}
prefix <- "processRepeats_Fustric"
cmd    <- with(metadataAssemblies, paste(processRepeatsPath, " -species fungi ", 
                                         paste(RepeatDBPath, "Final_mask_Fustric/*.full_mask.cat ", sep = ""), sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Finally, in order to feed these repeats into MAKER properly, we must separate out the complex repeats (more info on this below).  
**Create GFF3:**
```{r}
prefix <- "rmOutToGFF3_Fustric"
cmd    <- with(metadataAssemblies, 
               paste(rmOutToGFF3Path, " ", paste(RepeatDBPath, "Full_mask_Fustric/*.full_mask.out ", sep = ""),
                     " > ", paste(RepeatDBPath, "Full_mask_Fustric/", SppAbbrv, ".assembly.full_mask.out.gff3", sep = ""), sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Isolate complex repeats:**
```{r}
prefix <- "isolate_complex_Repeats_Fustric"
cmd    <- with(metadataAssemblies, 
               paste('grep -v -e "Satellite" -e ")n" -e "-rich"',
                     " ", paste(RepeatDBPath, "Full_mask_Fustric/", SppAbbrv, ".assembly.full_mask.out.gff3", sep = ""),
                     " > ", paste(RepeatDBPath, "Full_mask_Fustric/", SppAbbrv, ".assembly.full_mask.complex.gff3", sep = ""),
                     sep = ""))
            
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Reformat to work with MAKER:**   
**Note:**  
I can't seem to be able to generate the commands I want for this one. The commands are in the chunk below, and 
I passed them on the terminal in each directory manually.
```{r}
cmd <- with(metadataAssemblies,
            paste("cat ", paste(RepeatDBPath, "Full_mask_Fustric/", SppAbbrv, ".assembly.full_mask.complex.gff3 ", sep = ""),
                  " | ",
                  cat(paste("perl -ane '$id; if(!/^\\#/){@F = split(/\\t/, $_); chomp $F[-1];$id++; $F[-1] .= \"\\;ID=$id\"; $_ = join(\"\\t\", @F).\"\\n\"} print $_' ")),
                  " > ",
                  paste(RepeatDBPath, "Full_mask_Fustric/", SppAbbrv, ".assembly.full_mask.complex.reformat.gff3 ", sep = ""),
                  sep =""))
cmd
# cat Full_mask_Fustric/Fustric.assembly.full_mask.complex.gff3 | \
#  perl -ane '$id; if(!/^\#/){@F = split(/\t/, $_); chomp $F[-1];$id++; $F[-1] .= "\;ID=$id"; $_ = join("\t", @F)."\n"} print $_' \
#  > Full_mask_Fustric/Fustric.assembly.full_mask.complex.reformat.gff3
```
cat Fustric.assembly.full_mask.complex.gff3 | \\  
perl -ane '$id; if(!/^\#/){@F = split(/\t/, $_); chomp $F[-1];$id++; $F[-1] .= "\;ID=$id"; $_ = join("\t", @F)."\n"} print $_' \  
\> Fustric.assembly.full_mask.complex.reformat.gff3  

Now we have the prerequisite data for running MAKER. Add this to the data table:
```{r}
metadataAssemblies$makerFullMaskGFF3 <- paste(metadataAssemblies$RepeatDBPath, "Full_mask_Fustric/",
                                              metadataAssemblies$SppAbbrv, ".assembly.full_mask.complex.reformat.gff3", sep = "")
```

### 2. Annotation using Genemark-ES.  
Before we run MAKER, we can run Genemark-es, which requires no other input but the genome assemblies. This is supposed to be especially 
useful for fungal genomes. Once this is running, you can continue on with the chunk that gets the initial run of MAKER going, since it
doesn't depend on the genemark-es outputs for the first round of MAKER.  
It's a bit tricky to install the genemark-es due to dependencies, but works by installing one at a time all
dependencies using **cpanm**. Also, there was an error when trying to run it at first:  
  
/usr/bin/perl: symbol lookup error: /home/CFIA-ACIA/girouxeml/perl5/lib/perl5/x86_64-linux-thread-multi/auto/List/Util/Util.so: undefined symbol: Perl_xs_apiversion_bootcheck  
  
The fix was to remove the directory in _**~/perl5/lib/perl5/x86_64-linux-thread-multi/auto/List/Util/**_ and just 
placing it in my home directory (or if you want you can just delete it too):  
  
**_$ cd ~/perl5/lib/perl5/_**  
**_$ mv -f x86_64-linux-thread-multi_** **~**  
  
See: https://www.reddit.com/r/openbsd/comments/65efvo/openbsd_61_perl_problem_undefined_symbol_perl_xs/  
  

#### 2-A. Setting up paths to run Genemark-es in the species RepeatDB directory:
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[k], "genemark-es_round1", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$genemarkEsRnd1Path <- paste(metadataAssemblies$RepeatDBPath, "genemark-es_round1/", sep = "")
```

#### 2-B. Run The first round of Genemark-es in the maker_round1 dir:
```{r}
prefix <- "genemark-ES_rnd1_Fustric"
node   <- 8
cmd    <- with(metadataAssemblies,
               paste("cd ", genemarkEsRnd1Path, " && ",
                     " perl ", genemarkPath, " --ES --fungus --cores ", node, " --v ",
                     " --sequence ", FixedNewAssemName, sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 2-C. Rename the output file to include the organism name:
```{r}
cmd <- with(metadataAssemblies,
            paste("cd ", genemarkEsRnd1Path, " && ",
                  "mv output/gmhmm.mod output/", paste(SppAbbrv, "_GMES.mod", sep = ""), sep = ""))
sapply(cmd, function(x) system(x))
```

### 3. MAKER Round 1: Initial MAKER Analysis.  
MAKER is pretty easy to get going and relies on properly completed control files. In this first round, we will be
providing the data files for the repeat annotation (rm_gff), the transcriptome assembly (est), and extra protein files from
related species, which we have retrieved and saved in our References directory. We will also set the model_org to 'simple' 
so that only simple repeats are annotated (along with RepeatRunner).
```{r}
for(i in 1:length(metadataAssemblies$RepeatDBPath)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[i], "maker_round_1", sep = ""), showWarnings = TRUE, recursive = FALSE)
}
metadataAssemblies$maker_rnd1_path <- paste(metadataAssemblies$RepeatDBPath, "maker_round_1/", sep = "")
```

#### 3-A. Generate the MAKER control files.
```{r}
prefix <- "maker_rnd1_CTL_Fustric"
cmd    <- with(metadataAssemblies, paste("cd ", maker_rnd1_path, " && ", makerPath, " -CTL ", sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
metadataAssemblies$makerRnd1_CTL_optsPath <- paste(metadataAssemblies$maker_rnd1_path, "maker_opts.ctl", sep = "")
metadataAssemblies$makerRnd1_CTL_exePath  <- paste(metadataAssemblies$maker_rnd1_path, "maker_exe.ctl", sep = "")
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```
#### 3-B. Edit the MAKER control files.  
**The commands below will edit the altest, altest_gff, model_org, augustus_species, est2genome, protein2genome, and min_contig:**
```{r}
cmd <- with(metadataAssemblies,
            paste("/opt/rocks/bin/replace",
                  " altest= ", paste("altest=", cdnaFuspoaPath, sep = ""),
                  " altest_gff= ", paste("altest_gff=", gff3FuspoaPath, sep = ""),
                  " model_org=all ", "model_org=simple ",
                  " augustus_species= ", "augustus_species=fusarium_graminearum ",
                  " est2genome=0 ", "est2genome=1 ",
                  " protein2genome=0 ", "protein2genome=1 ",
                  " min_contig=1 ", "min_contig=1000 ",
                  " -- ", makerRnd1_CTL_optsPath,
                  sep = ""))

sapply(cmd, function(x) system(x))
```

**Edit the maker_opts.ctl genome paths, and protein paths and rm_gff (repeat modeler) paths to point to desired files:**
```{r}
# Edit Genome paths:
newPattern <- paste("genome=", metadataAssemblies$FixedNewAssemName, sep = "")
pattern    <- "genome= "

for(i in 1:nrow(metadataAssemblies)){
    makerOptsGenRepl(metadataAssemblies$makerRnd1_CTL_optsPath[i], pattern, newPattern[i])}

# Edit protein path pattern, using this function to make sure "protein= " is edited, and not
# "repeat_protein=/home/CFIA-ACIA/girouxeml/prog/maker/data/te_proteins.fasta".
pattern    <- "protein= "
newPattern <- paste("protein=", pepFuspoa2Path, ",", pepFusavePath, ",", pepFuslangPath, 
                    ",", pepFuspoaPath, ",", pepFusfujPath, ",", pepFusgramPath, 
                    ",", pepFusmangPath, ",", pepFusoxysPath, ",", pepFusprolPath, 
                    ",", pepFuspseuPath, ",", pepFussolPath, ",", pepFusvertPath,
                    sep = "")

for(i in 1:nrow(metadataAssemblies)){
    makerOptsGenRepl(metadataAssemblies$makerRnd1_CTL_optsPath[i], pattern, newPattern)}

# Edit repeat modeler, "rm_gff= ", path:
pattern    <- "rm_gff= "
newPattern <- paste("rm_gff=", metadataAssemblies$makerFullMaskGFF3, sep = "")

for(i in 1:nrow(metadataAssemblies)){
    makerOptsGenRepl(metadataAssemblies$makerRnd1_CTL_optsPath[i], pattern, newPattern[i])}
```

#### 3-C. Run MAKER.  
Given MAKER will be using BLAST to align transcripts and proteins to the genome, this will take at 
least a couple days with 12 cores. Speed is a product of the resources you allow (more cores == faster) 
and the assembly quality (smaller, less contiguous scaffolds == longer).

Need to use our conda env for the following, or we will have issues with "Fstack_protector_strong"
$ conda activate makerenv   
$ cd /Fusarium_tricinctum/Fustric_RepeatDB/maker_round_1/   
$ maker -base Fustric_rnd1 maker_opts.ctl maker_bopts.ctl maker_exe.ctl -cpus 15 2>&1 | tee maker.log
```{r}
# prefix <- "maker_round_1_run_Fustric"
# node   <- 12
# cmd    <- with(metadataAssemblies,
#                paste("cd ", maker_rnd1_path, " && perl ", makerPath, 
#                      " -base ", paste(SppAbbrv, "_rnd1", sep = ""),
#                      " ", paste(maker_rnd1_path, "maker_opts.ctl", sep = ""),
#                      " ", paste(maker_rnd1_path, "maker_bopts.ctl", sep = ""),
#                      " ", paste(maker_rnd1_path, "maker_exe.ctl", sep = ""),
#                      " -cpus 12 ", sep = ""))
# 
# suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```

**To remove the output files after you are done:**
```{r}
# RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 3-D. Assemble the MAKER GFF and FASTA outputs.
```{r}
metadataAssemblies$makerRnd1_OutputPath <- paste(metadataAssemblies$maker_rnd1_path, metadataAssemblies$SppAbbrv, "_rnd1.maker.output/", sep = "")

# GFF with the sequences:
prefix <- "mergeGFF3_maker_all_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", makerRnd1_OutputPath, " && ", makerGFF3merge, " -s -d ",
                     paste(makerRnd1_OutputPath, SppAbbrv, "_rnd1_master_datastore_index.log", sep = ""),
                     " > ", paste(makerRnd1_OutputPath, SppAbbrv, "_rnd1.all.maker.gff3", sep = ""), sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

# Collect all the fastas:
prefix2 <- "fastaMerge_maker_Fustric"
cmd2    <- with(metadataAssemblies,
                paste("cd ", makerRnd1_OutputPath, " && ", makerFastamerge, " -d ",
                      paste(makerRnd1_OutputPath, SppAbbrv, "_rnd1_master_datastore_index.log", sep = ""),
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)

# GFF without the sequences:
prefix3 <- "mergeGFF3_maker_all_noseq_Fustric" 
cmd3    <- with(metadataAssemblies,
                paste("cd ", makerRnd1_OutputPath, " && ", makerGFF3merge, " -n -s -d ",
                      paste(makerRnd1_OutputPath, SppAbbrv, "_rnd1_master_datastore_index.log", sep = ""),
                      " > ", paste(makerRnd1_OutputPath, SppAbbrv, "_rnd1.all.maker.noseq.gff3", sep = ""),
                     sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd3, prefix3, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2); RemoveQsubTempFiles(sharedPathAn, prefix3)
```

### 4. Training Gene Prediction Software SNAP and Augustus.  
Besides mapping the empirical transcript and protein evidence to the reference genome and repeat annotation 
(not much of this in our example, given we've done so much up front), the most important product of this MAKER 
run is the gene models. These are used for training gene prediction software like augustus and snap.  
  
#### 4-A. Training SNAP.  
SNAP is pretty quick and easy to train. Issuing the following commands will perform the training. It's best 
to put some thought into what kind of gene models you use from MAKER. Ideally we would like to have sufficient 
evidence to filter to use models with an AED of 0.25 or better and a length of 50 or more amino acids 
(maker2zff -x 0.25 -l 50), which helps get rid of junky models, but we do not have enough evidence to support this 
for our data (no EST or transcript data for our specific organism), and it will result in nothing being returned. 
But we can specify all filter parameters so that at least we can filter based on AED and length, and set all other options
dependant on EST evidence to 0, which is still better than setting it to no filters (-n).  
  
We start with exporting 'confident' gene models from round 1 of MAKER for use in retraining SNAP. We'll generate some summary stats
of the models, and then collecting the training sequences and generating training parameters for BUSCO to train Augustus.  
  
**Export confident gene models from round 1 of MAKER using maker2zff:**  
For determining which genes are High Confidence for Retraining:
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$maker_rnd1_path[k], "snap", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$snapMakerRnd1Path <- paste(metadataAssemblies$maker_rnd1_path, "snap/", sep = "")

for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$snapMakerRnd1Path[k], "round1", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$snapRnd1MakerRnd1Path <- paste(metadataAssemblies$snapMakerRnd1Path, "round1/", sep = "")

prefix <- "snapMaker2zff_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", snapRnd1MakerRnd1Path, " && ", maker2zffPath, " -x 0.25 -l 50 -c 0 -e 0 -o 0 -d ", 
                     paste(makerRnd1_OutputPath, SppAbbrv, "_rnd1_master_datastore_index.log", sep = ""), sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Take a quick look at number of seqs in genome.dna file:**
```{r}
cmd <- with(metadataAssemblies, paste("cd ", snapRnd1MakerRnd1Path, " && ", 'grep ">" genome.dna | wc -l ', sep = ""))
sapply(cmd, function(x) system(x))

# 20 contigs...?
cmd <- with(metadataAssemblies, file.rename(paste(snapRnd1MakerRnd1Path, "genome.dna", sep = ""),
                                            paste(snapRnd1MakerRnd1Path, SppAbbrv, "_rnd1.zff.dna", sep = "")))
sapply(cmd, function(x) system(x))

cmd <- with(metadataAssemblies, file.rename(paste(snapRnd1MakerRnd1Path, "genome.ann", sep = ""),
                                            paste(snapRnd1MakerRnd1Path, SppAbbrv, "_rnd1.zff.ann", sep = "")))
sapply(cmd, function(x) system(x))
```

**Gather some stats and validate using fathom:**
```{r}
prefix <- "fathom_MakerRnd1_stats_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", snapRnd1MakerRnd1Path, " && ",
                     fathomPath, " ", paste(SppAbbrv, "_rnd1.zff.ann", sep = ""),
                     " ", paste(SppAbbrv, "_rnd1.zff.dna", sep = ""),
                     " -gene-stats > gene-stats.log 2>&1 ", sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

prefix2 <- "fathom_MakerRnd1_validate_Fustric"
cmd2    <- with(metadataAssemblies, 
                paste("cd ", snapRnd1MakerRnd1Path, " && ", 
                      fathomPath, " ", paste(SppAbbrv, "_rnd1.zff.ann", sep = ""),
                      " ", paste(SppAbbrv, "_rnd1.zff.dna", sep = ""),
                      " -validate > validate.log 2>&1 ", sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2)
```

**Collect the training sequences and annotations, plus 1000 surrounding bp for training using fathom:**
```{r}
prefix <- "fathom_MakerRnd1_categorize_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", snapRnd1MakerRnd1Path, " && ",
                     fathomPath, " ", paste(SppAbbrv, "_rnd1.zff.ann", sep = ""),
                     " ", paste(SppAbbrv, "_rnd1.zff.dna", sep = ""),
                     " -categorize 1000 > categorize.log 2>&1 ", sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

Wait for the chunk above to complete before running this next one. It depends on the output from the 
previous as input.
```{r}
prefix2 <- "fathom_MakerRnd1_export_Fustric"
cmd2    <- with(metadataAssemblies,
                paste("cd ", snapRnd1MakerRnd1Path, " && ",
                      fathomPath, " uni.ann uni.dna -export 1000 -plus > uni-plus.log 2>&1 ", sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2)
```

**Create the training parameters:**
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$snapMakerRnd1Path[k], "round1/params", sep = ""),
               showWarnings = TRUE, recursive = FALSE)}

prefix <- "forge_MakerRnd1_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", paste(snapRnd1MakerRnd1Path, "params/", sep = ""), " && ",
                     forgePath, " ../export.ann ../export.dna > ../forge.log 2>&1 ", sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Assemble the HMM:**
```{r}
prefix <- "hmmAssembler_MakerRnd1_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", snapRnd1MakerRnd1Path, " && ",
                     hmmAssemPath, " ", paste(SppAbbrv, "_rnd1.zff", sep = ""),
                     " params > ", paste(SppAbbrv, "_rnd1.zff.hmm", sep = ""), sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 4-B. Training Augustus using BUSCO.  
**Augustus**  
Training Augustus is a more laborious process. Luckily, the recent release of BUSCO provides a nice pipeline 
for performing the training, while giving you an idea of how good your annotation already is (If you don't want 
to go this route, there are scripts provided with Augustus to perform the training). First, the Parallel::ForkManager 
module for Perl is required to run BUSCO with more than one core. You can easily install it before the first time 
you use BUSCO by running:   
  
_**$ sudo apt-get install libparallel-forkmanager-perl**_  
_**$ cpanm Parallel::ForkManager**_  
  
**Excise regions with mRNA annotations from MAKER round 1**    
This probably isn't an ideal training environment, but appears to work well. First, we must put together training 
sequences using the gene models we created in our first run of MAKER. We do this by issuing the following command 
to excise the regions that contain mRNA annotations based on our initial MAKER run (with 1000bp on each side).
```{r}
prefix <- "maker-augustus_get_mRNA_regions_Rnd1_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", snapRnd1MakerRnd1Path, " && ",
                     paste(' awk -v OFS="\t" \'{ if ($3 == "mRNA") print $1, $4, $5 }\'', sep = ""),
                     " ", paste(makerRnd1_OutputPath, SppAbbrv, "_rnd1.all.maker.noseq.gff3", sep = ""),
                     " | \ ",
                     paste('awk -v OFS="\t" \'{ if ($2 < 1000) print $1, "0", $3+1000; else print $1, $2-1000, $3+1000 }\'', sep = ""),
                     " | \ ",
                     bedToolsPath, " getfasta -fi ", FixedNewAssemName, " -bed - -fo ",
                     paste(SppAbbrv, "_rnd1.all.maker.transcripts1000.fasta", sep = "")))

# Note, open the bash scripts, and edit so put the "\t" back in.
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

There are some important things to note based on this approach. First is that you will likely get warnings from 
BEDtools that certain coordinates could not be used to extract FASTA sequences. This is because the end coordinate 
of a transcript plus 1000 bp is beyond the total length of a given scaffold. This script does account for transcripts 
being within the beginning 1000bp of the scaffold, but there was no easy way to do the same with transcripts within 
the last 1000bp of the scaffold. This is okay, however, as we still end up with sequences from thousands of gene models 
and BUSCO will only be searching for a small subset of genes itself.  
  
While we've only provided sequences from regions likely to contain genes, we've totally eliminated any existing annotation 
data about the starts/stops of gene elements. Augustus would normally use this as part of the training process. However, 
BUSCO will essentially do a reannotation of these regions using BLAST and built-in HMMs for a set of conserved genes 
(hundreds to thousands). This has the effect of recreating some version of our gene models for these conserved genes. We 
then leverage the internal training that BUSCO can perform (the --long argument) to optimize the HMM search model to train 
Augustus and produce a trained HMM for MAKER. Here is the command we use to perform the Augustus training inside BUSCO.  
  
In this case, we are using the Sordariomyceta set of conserved genes (N=3725 genes)(the Fungi set of conserved genes is N = 290 genes), 
so BUSCO will try to identify those genes using BLAST and an initial HMM model for each that comes stocked within the BUSCO database. 
We specify the -m genome option since we are giving BUSCO regions that include more than just transcripts. The initial HMM model we'll use is the 
*Fusarium graminearum* one (-sp fusarium_graminearum), which is a reasonably close species. Finally, the --long 
option tells BUSCO to use the initial gene models it creates to optimize the HMM settings of the raw human HMM, thus training 
it for our use on Fus3. We can have this run in parallel on several cores, but it will still likely take days, so be patient. 
Note, some of the options are kept in the config.ini file in the busco/config/ directory. Also, if during the run it keeps 
crashing with segmentation fault, core dumped, it may be an error with the threading. Try running on just one core. It may 
take a lot longer though, but at least it's still moving along.  

**Run BUSCO:**
```{r}
prefix <- "Busco_MakerRnd1_Fustric"
node   <- 15
cmd    <- with(metadataAssemblies,
               paste("cd ", snapRnd1MakerRnd1Path, " && ",
                     "python ", buscoPath, " -i ", paste(SppAbbrv, "_rnd1.all.maker.transcripts1000.fasta", sep = ""),
                     " -o ", paste(SppAbbrv, "_rnd1_maker", sep = ""), " -c ", node,
                     " -l ", buscoFunDataSet, " -m genome -sp fusarium_graminearum -z --augustus_parameters='--progress=true'",
                     # " --long --restart",
                     " --long ",
                     " -f ", # If forced to restart and there is a problem with the previous run and you shouldn't use restart.
                     sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
# Note, if the run fails, restart the run with the following "--restart", it will continue from where it left off. However, check the 
# config.ini file in ~/prog/busco/config/ to make sure that restart is set to True, and that you adjust to the correct cpus number.
```
I should have used the BUSCO set sordariomyceta_odb9 instead of Fungi, see the following paper for reference:   
https://www.frontiersin.org/articles/10.3389/fmicb.2018.01158/full   


**Run BUSCO:**
```{r}
prefix <- "Busco_MakerRnd1_Fustric_Sordario"
node   <- 15
cmd    <- with(metadataAssemblies,
               paste("cd ", snapRnd1MakerRnd1Path, " && ",
                     "python ", buscoPath, " -i ", paste(SppAbbrv, "_rnd1.all.maker.transcripts1000.fasta", sep = ""),
                     " -o ", paste(SppAbbrv, "_rnd1_maker", sep = ""), " -c ", node,
                     " -l ", buscoSorDataSet, " -m genome -sp fusarium_graminearum -z --augustus_parameters='--progress=true'",
                     # " --long --restart",
                     " --long ",
                     " -f ", # If forced to restart and there is a problem with the previous run and you shouldn't use restart.
                     sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
# Note, if the run fails, restart the run with the following "--restart", it will continue from where it left off. However, check the 
# config.ini file in ~/prog/busco/config/ to make sure that restart is set to True, and that you adjust to the correct cpus number.
```


**To remove the output files after you are done:**
```{r}
# system("/opt/gridengine/bin/linux-x64/qstat") # Remove qsub temp when qstat returns nothing.
RemoveQsubTempFiles(sharedPathAn, prefix)
cmd <- with(metadataAssemblies, paste("cd ", snapRnd1MakerRnd1Path, " && rm -rf tmp ", sep = ""))
sapply(cmd, function(x) system(x))
cmd <- with(metadataAssemblies, paste("cd ", snapRnd1MakerRnd1Path, " && rm core.* ", sep = ""))
sapply(cmd, function(x) system(x))
```

Once BUSCO is complete, it will give you an idea of how complete your annotation is (though be cautious, 
because we haven't filtered away known alternative transcripts that will be binned as duplicates). We need 
to do some post-processing of the HMM models to get them ready for MAKER. First, we'll rename the files within
/maker_round_1/snap/round1/run_Lsue_rnd1_maker/augustus_output/retraining_parameters. 
  
**Rename output files from the BUSCO run:**
```{r}
metadataAssemblies$BuscoOut <- paste(metadataAssemblies$snapRnd1MakerRnd1Path, "run_", metadataAssemblies$SppAbbrv, "_rnd1_maker_Sordario/", sep = "")
metadataAssemblies$BuscoRetrainParam <- paste(metadataAssemblies$BuscoOut, "augustus_output/retraining_parameters/", sep = "")

fl <- list.files(metadataAssemblies$BuscoRetrainParam, full.names = TRUE)

# Rather than develop a script, I went to each dir and used rename on the files:
# ~/perl5/bin/rename 's/BUSCO_Fustric_rnd1_maker_Sordario_2109611257/Fusarium_tricinctum/g' *
```

**Rename the files cited within the output files from the BUSCO run:**  
We also need to rename the files cited within certain HMM configuration files:
```{r}
# sed -i 's/BUSCO_Fustric_rnd1_maker_Sordario_2109611257/Fusarium_tricinctum/g' Fusarium_tricinctum_parameters.cfg
```

**Add the newly-trained Augustus retraining parameter files to the Augustus species library directory:**  
Finally, we must copy these into the $AUGUSTUS_CONFIG_PATH species HMM location so they are accessible 
by Augustus and MAKER. I had put the $AUGUSTUS_CONFIG_PATH in the .bash_profile, after copying the 
directory from the isilon MAKER pipeline, for MAKER version 2.10. This is because I do not have 
write access to the isilon directory, but I do have it for my local directory.  
  
Do this from within the dir /maker_round_1/snap/round1/run_Lsue_rnd1_maker/augustus_output/retraining_parameters/
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste("/isilon/cfia-ottawa-fallowfield/users/girouxeml/prog/augustus/config/species/", 
                     metadataAssemblies$ScientificName[k], sep = ""),
               showWarnings = TRUE, recursive = FALSE)
}

cmd <- with(metadataAssemblies,
            paste("cd ", BuscoRetrainParam, " && cp ", ScientificName, "*", 
                  " /isilon/cfia-ottawa-fallowfield/users/girouxeml/prog/augustus/config/species/", ScientificName, sep = ""))
sapply(cmd, function(x) system(x))
```

### 5. MAKER round 2 with ab initio Gene Predictors.  
Now let's run a second round of MAKER, but this time we will have Genemark-es, SNAP and Augustus run 
within MAKER to help create more sound gene models. MAKER will use the annotations from these two 
prediction programs when constructing its models. Before running, let's first recycle the mapping of 
empicial evidence we have from the first MAKER round, so we don't have to perform all the BLASTs, etc. again.

#### 5-A. Recycle the mapping and empirical evidence obtained from the first round with MAKER:
```{r}
# Transcript alignments - our file will be empty, as we have no est or transcript evidence, so may need to fix this:
prefix <- "maker_rnd2_transcript_evidence_fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", makerRnd1_OutputPath, " && ",
                     paste(" awk '{ if ", '($2 == "est2genome") print $0 }', "'", sep = ""),
                     " ", paste(SppAbbrv, "_rnd1.all.maker.noseq.gff3", sep = ""), 
                     " > ", paste(SppAbbrv, "_rnd1.all.maker.est2genome.gff3", sep = ""),
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

# Protein alignments:
prefix2 <- "maker_rnd2_protein_evidence_fustric"
cmd     <- with(metadataAssemblies,
                paste("cd ", makerRnd1_OutputPath, " && ",
                      paste(" awk '{ if ", '($2 == "protein2genome") print $0 }', "'", sep = ""),
                      " ", paste(SppAbbrv, "_rnd1.all.maker.noseq.gff3", sep = ""), 
                      " > ", paste(SppAbbrv, "_rnd1.all.maker.protein2genome.gff3", sep = ""),
                      sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix2, suffix)

# Repeat alignments:
prefix3 <- "maker_rnd2_repeat_evidence_fustric"
cmd     <- with(metadataAssemblies,
                paste("cd ", makerRnd1_OutputPath, " && ",
                      paste(" awk '{ if ", '($2 ~ "repeat") print $0 }', "'", sep = ""),
                      " ", paste(SppAbbrv, "_rnd1.all.maker.noseq.gff3", sep = ""), 
                      " > ", paste(SppAbbrv, "_rnd1.all.maker.repeats.gff3", sep = ""),
                      sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix3, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2); RemoveQsubTempFiles(sharedPathAn, prefix3)
```

**Delete the unnecessary tmp files created by BUSCO:**
```{r}
# prefix <- "delete_dir"
# cmd <- with(metadataAssemblies, paste("cd ", RepeatDBPath, " && rm -rf maker_round_2 ", sep = ""))
# suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Now go up to the RepeatDB dir, and create a dir for the next round of MAKER:**
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[k], "maker_round_2", sep = ""), 
               showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$maker_rnd2_path <- paste(metadataAssemblies$RepeatDBPath, "maker_round_2/", sep = "")
```

#### 5-B. Generate the MAKER round 2 control files:  
```{r}
prefix <- "maker_rnd2_CTL"
cmd    <- with(metadataAssemblies, paste("cd ", maker_rnd2_path, " && ", makerPath, " -CTL ", sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
metadataAssemblies$makerRnd2_CTL_optsPath <- paste(metadataAssemblies$maker_rnd2_path, "maker_opts.ctl", sep = "")
metadataAssemblies$makerRnd2_CTL_exePath  <- paste(metadataAssemblies$maker_rnd2_path, "maker_exe.ctl", sep = "")
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```
  
#### 5-C. Edit the MAKER round 2 control files.  
  
Modify the control file, removing the FASTA sequences files to map and replacing them with the GFFs (est_gff, 
protein_gff, and rm_gff, respectively. We can also specify the path to the Genemark-es .mod, SNAP HMM and the 
species name for Augustus, so that these gene prediction programs are run. We will also switch est2genome and 
protein2genome to 0 so that gene predictions are based on the Augustus and SNAP gene models. Unfortunately we 
do not have est or transcript evidence to use, so instead I used the gff3 of the botrytis cinerea model, which 
may be a mistake. I set est2genome=0. I also installed the trnascan program locally, following the instructions 
on: https://iamphioxus.org/2016/01/20/installing-trnascan-se-and-snoscan/ and added the following to my .bash_profile:
$ export Path="/home/CFIA-ACIA/girouxeml/prog/tRNAscan-SE/bin:$PATH"
$ PERL5LIB=~prog/tRNAscan-SE/bin

Note: If working in a conda environment, such as makerenv that I've created, create a sym link that will
point to where I am storing the augustus/config/species directories. This is bacause conda writes everything to my home 
directory, which is limited for space, so I store the databases where I have more space:   
$ ln -s /isilon/cfia-ottawa-fallowfield/users/girouxeml/prog/augustus/config/species /home/CFIA-ACIA/girouxeml/prog/miniconda/envs/makerenv/config/species
```{r}
cmd <- with(metadataAssemblies,
            paste("/opt/rocks/bin/replace ",
                  " altest= ", paste("altest=", cdnaFuspoaPath, sep = ""),
                  " altest_gff= ", paste("altest_gff=", gff3FuspoaPath, sep = ""),
                  " protein_gff= ", paste("protein_gff=", makerRnd1_OutputPath, SppAbbrv, "_rnd1.all.maker.protein2genome.gff3", sep = ""),
                  " model_org=all model_org=# ",
                  " repeat_protein= repeat_protein=# ",
                  " rm_gff= ", paste("rm_gff=", makerRnd1_OutputPath, SppAbbrv, "_rnd1.all.maker.repeats.gff3", sep = ""),
                  " snaphmm= ", paste("snaphmm=", snapRnd1MakerRnd1Path, SppAbbrv, "_rnd1.zff.hmm", sep = ""),
                  " gmhmm= ", paste("gmhmm=", genemarkEsRnd1Path, "output/", SppAbbrv, "_GMES.mod", sep = ""),
                  " augustus_species= ", paste("augustus_species=", ScientificName, " ", sep = ""),
                  " trna=0 trna=1 ",
                  " max_dna_len=100000 max_dna_len=300000 ",
                  " -- ", makerRnd2_CTL_optsPath,
                  sep = ""))
sapply(cmd, function(x) system(x))

# I also modified the maker_exe.ctl with the local path to tRNAscan:
pattern    <- paste("tRNAscan-SE=", pathtRNA_scan, sep = "")
newPattern <- paste("tRNAscan-SE=", pathtRNA_scan, sep = "")
cmd        <- with(metadataAssemblies, paste("/opt/rocks/bin/replace ", pattern, " ", newPattern,
                                             " -- ", makerRnd2_CTL_exePath, sep = ""))
sapply(cmd, function(x) system(x))

# Edit Genome paths:
pattern    <- "genome= "
newPattern <- paste("genome=", metadataAssemblies$FixedNewAssemName, sep = "")
for(i in 1:nrow(metadataAssemblies)){
    makerOptsGenRepl(metadataAssemblies$makerRnd2_CTL_optsPath[i], pattern, newPattern[i])}
```

#### 5-D. Run MAKER round 2.  
**Then we can run MAKER, substituting this new control file, and summarize the output, as we did before.**
```{r}
prefix <- "maker_round_2_run_fustric"
node   <- 20 # qlogin with pe smp 20, run in conda makerenv
cmd    <- with(metadataAssemblies,
               paste("cd ", maker_rnd2_path, " && maker ", # For this, I'm running in conda makerenv.
                     " -base ", paste(SppAbbrv, "_rnd2", sep = ""),
                     " ", paste(maker_rnd2_path, "maker_opts.ctl", sep = ""),
                     " ", paste(maker_rnd2_path, "maker_bopts.ctl", sep = ""),
                     " ", paste(maker_rnd2_path, "maker_exe.ctl", sep = ""),
                     " -cpus ", node, " 2>&1 | tee makerRnd2.log", sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 5-E. Assemble the MAKER round 2 GFF and FASTA outputs.  
**We conclude by assembling together the GFF and FASTA outputs.**
```{r}
metadataAssemblies$makerRnd2_OutputPath <- paste(metadataAssemblies$maker_rnd2_path, 
                                                 metadataAssemblies$SppAbbrv, "_rnd2.maker.output/", sep = "")
# GFF with the sequences:
prefix  <- "round2_mergeGFF3_maker_all_Fustric"
cmd     <- with(metadataAssemblies,
                paste(" conda activate makerenv && ", # test to see if this can be done with qsub  
                      " cd ", makerRnd2_OutputPath, 
                      " && ", makerGFF3merge, " -s -d ",
                      paste(makerRnd2_OutputPath, SppAbbrv, "_rnd2_master_datastore_index.log", sep = ""),
                      " > ", paste(makerRnd2_OutputPath, SppAbbrv, "_rnd2.all.maker.gff3", sep = ""), 
                      " && conda deactivate ", # test to make sure this works with qsub
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

# Collect all the fastas:
prefix2 <- "round2_fastaMerge_maker_Fustric"
cmd     <- with(metadataAssemblies,
                paste("conda activate makerenv && ", # test to see if this can be done with qsub
                      "cd ", makerRnd2_OutputPath, " && ", makerFastamerge, " -d ",
                      paste(makerRnd2_OutputPath, SppAbbrv, "_rnd2_master_datastore_index.log", sep = ""),
                      " && conda deactivate ", # test to make sure this works with qsub
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix2, suffix)

# GFF without the sequences:
prefix3 <- "round2_mergeGFF3_maker_all_noseq_Fustric"
cmd     <- with(metadataAssemblies,
                paste("conda activate makerenv && ", # test to see if this can be done with qsub
                      "cd ", makerRnd2_OutputPath, " && ", makerGFF3merge, " -n -s -d ",
                      paste(makerRnd2_OutputPath, SppAbbrv, "_rnd2_master_datastore_index.log", sep = ""),
                      " > ", paste(makerRnd2_OutputPath, SppAbbrv, "_rnd2.all.maker.noseq.gff3", sep = ""),
                      " && conda deactivate ", # test to make sure this works with qsub
                      sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix3, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2); RemoveQsubTempFiles(sharedPathAn, prefix3)
```

**Export confident gene models from round 1 of MAKER using maker2zff:**  
For determining which genes are High Confidence for Retraining:
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$maker_rnd2_path[k], "snap", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$snapMakerRnd2Path <- paste(metadataAssemblies$maker_rnd2_path, "snap/", sep = "")

for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$snapMakerRnd2Path[k], "round1", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$snapRnd1MakerRnd2Path <- paste(metadataAssemblies$snapMakerRnd2Path, "round1/", sep = "")

prefix <- "round2_snapMaker2zff_Fustric"
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", snapRnd1MakerRnd2Path, " && ", maker2zffPath, " -x 0.25 -l 50 -c 0 -e 0 -o 0 -d ", 
                     paste(makerRnd2_OutputPath, SppAbbrv, "_rnd2_master_datastore_index.log", sep = ""),
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Take a quick look at number of sequences in the genome.dna file:**
```{r}
cmd <- with(metadataAssemblies, paste("cd ", snapRnd1MakerRnd2Path, " && ", 'grep ">" genome.dna | wc -l ', sep = ""))
sapply(cmd, function(x) system(x))

cmd <- with(metadataAssemblies, file.rename(paste(snapRnd1MakerRnd2Path, "genome.dna", sep = ""),
                                            paste(snapRnd1MakerRnd2Path, SppAbbrv, "_rnd2.zff.dna", sep = "")))
sapply(cmd, function(x) system(x))

cmd <- with(metadataAssemblies, file.rename(paste(snapRnd1MakerRnd2Path, "genome.ann", sep = ""),
                                            paste(snapRnd1MakerRnd2Path, SppAbbrv, "_rnd2.zff.ann", sep = "")))
sapply(cmd, function(x) system(x))
```

**Gather some stats and validate:**
```{r}
prefix  <- "round2_fathom_stats_Fustric"
cmd     <- with(metadataAssemblies, paste("conda activate makerenv && ",
                                          "cd ", snapRnd1MakerRnd2Path, " && ", fathomPath, " ",
                                          paste(SppAbbrv, "_rnd2.zff.ann", sep = ""), " ", paste(SppAbbrv, "_rnd2.zff.dna", sep = ""),
                                          " -gene-stats > gene-stats.log 2>&1 ",
                                          " && conda deactivate ",
                                          sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

prefix2 <- "round2_fathom_validate_Fustric"
cmd2    <- with(metadataAssemblies, paste("conda activate makerenv && ",
                                          "cd ", snapRnd1MakerRnd2Path, " && ", fathomPath, " ",
                                          paste(SppAbbrv, "_rnd2.zff.ann", sep = ""), " ", paste(SppAbbrv, "_rnd2.zff.dna", sep = ""),
                                          " -validate > validate.log 2>&1 ",
                                          " && conda deactivate ",
                                          sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2)
```

**Collect the training sequences and annotations, plus 1000 surrounding bp for training**
```{r}
prefix <- "round2_fathom_categorize_Fustric"
cmd    <- with(metadataAssemblies, paste("conda activate makerenv && ",
                                         "cd ", snapRnd1MakerRnd2Path, " && ", fathomPath, " ",
                                         paste(SppAbbrv, "_rnd2.zff.ann", sep = ""), " ", paste(SppAbbrv, "_rnd2.zff.dna", sep = ""),
                                         " -categorize 1000 > categorize.log 2>&1 ",
                                         " && conda deactivate ",
                                         sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Wait for the categorize run to finish before doing the export, which is next.**
```{r}
prefix2 <- "round2_fathom_export_Fustric"
cmd2    <- with(metadataAssemblies, 
                paste("conda activate makerenv && ",
                      "cd ", snapRnd1MakerRnd2Path, " && ",
                      fathomPath, " uni.ann uni.dna -export 1000 -plus > uni-plus.log 2>&1 ",
                      " && conda deactivate ",
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix2)
```

**Create the training parameters for SNAP:**
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$snapRnd1MakerRnd2Path[k], "params", sep = ""),
               showWarnings = TRUE, recursive = FALSE)}

prefix <- "round2_forge_Fustric"
cmd <- with(metadataAssemblies,
            paste("conda activate makerenv && ",
                  "cd ", paste(snapRnd1MakerRnd2Path, "params/", sep = ""), " && ",
                  forgePath, " ../export.ann ../export.dna > ../forge.log 2>&1 ",
                  " && conda deactivate ",
                  sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Assemble the HMM:**
```{r}
prefix <- "round2_hmmAssembler_Maker_Fustric"
cmd <- with(metadataAssemblies,
            paste("conda activate makerenv && ",
                  "cd ", snapRnd1MakerRnd2Path, " && ",
                  hmmAssemPath, " ", paste(SppAbbrv, "_rnd2.zff", sep = ""),
                  " params > ", paste(SppAbbrv, "_rnd2.zff.hmm", sep = ""),
                  " && conda deactivate ",
                  sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

### We've now completed the 2nd round of MAKER.   
#### 5-F. Evaluate the gene models.  
We've now completed the 2nd round of MAKER and should evaluate the gene models we have so far:  

**i. Count the number of gene models and the gene lengths:**    
I manually recorded the output values to a summary BUSCO excel file. Welcome to ideas on how to capture the output.
```{r}
cmd <- with(metadataAssemblies, 
            paste("cd ", makerRnd2_OutputPath, " && ", " cat ", SppAbbrv, "_rnd2.all.maker.gff3 ", 
                  " | awk '{ if ($3 ==", '"gene") print $0}', 
                  "' | awk '{ sum += ($5 - $4) } END { print NR, sum / NR }'", sep = ""))

sapply(cmd, function(x) system(x))
```

**ii. Visualize the AED distribution:**  
AED ranges from 0 to 1 and quantifies the confidence in a gene model 
based on empirical evidence. Basically, the lower the AED, the better a gene model is likely to be. Ideally, 
95% or more of the gene models will have an AED of 0.5 or better in the case of good assemblies. You can use 
this AED_cdf_generator.pl script to help with this.  
I have just recorded the fraction at 0.5 into my summary excel table. Open to ideas on how to capture it better. 
```{r}
cmd <- with(metadataAssemblies, 
            paste("cd ", makerRnd2_OutputPath, " && ", " perl ", aedCDFgenePath, 
                  " -b 0.025 ", SppAbbrv, "_rnd2.all.maker.gff3", sep = "")) 
sapply(cmd, function(x) system(x))
```

**iii. Run BUSCO using the Augustus species HMM to look at the results:**  
We can run BUSCO using the Augustus species HMM we obtained, without having to train Augustus. Only include the 
transcript sequences, and not the 1000 bp on each side, and be sure to take the best (i.e., longest) transcript for
each gene so we aren't artificially seeding duplicates. We could also run it on the best protein sequence per gene
instead. 
```{r}
prefix <- "round2_eval_with_Busco_Fustric"
node   <- 1
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", makerRnd2_OutputPath, " && ",
                     "python ", buscoPath, " -i ", paste(SppAbbrv, "_rnd2.all.maker.transcripts.fasta", sep = ""),
                     " -o annotation_eval ", " -c ", node,
                     " -l ", buscoSorDataSet, " -m transcriptome -sp ", ScientificName, " -z --augustus_parameters='--progress=true'",
                     # " --restart ", 
                     # " --force ",
                     " && conda deactivate ",
                     sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
# Note, if the run fails, restart the run with the following "--restart", it will continue from where it left off. However, check the 
# config.ini file in ~/prog/busco/config/ to make sure that restart is set to True, and that you adjust to the correct cpus number.
```
# BUSCO version is: 3.0.2
# The lineage dataset is: sordariomyceta_odb9 (Creation date: 2016-02-13, number of species: 30, number of BUSCOs: 3725)
# To reproduce this run: python /home/CFIA-ACIA/girouxeml/prog/busco/scripts/run_BUSCO.py -i Fustric_rnd2.all.maker.transcripts.fasta -o annotation_eval -l /isilon/cfia-ottawa-fallowfield/users/girouxeml/prog/busco_datasets/sordariomyceta_odb9/ -m transcriptome -c 1 -z
#
# Summarized benchmarking in BUSCO notation for file Fustric_rnd2.all.maker.transcripts.fasta
# BUSCO was run in mode: transcriptome

        C:95.2%[S:95.0%,D:0.2%],F:3.0%,M:1.8%,n:3725

        3546    Complete BUSCOs (C)
        3537    Complete and single-copy BUSCOs (S)
        9       Complete and duplicated BUSCOs (D)
        113     Fragmented BUSCOs (F)
        66      Missing BUSCOs (M)
        3725    Total BUSCO groups searched

### 6. Training Augustus - Round 2  
We won't be using BUSCO to train Augustus the second time since as there is a bias in the models it
produces for conserved genes that tend to be short and intron poor. To avoid this bias in the second round, we use a broad selection 
of gene models instead. After using the maker2zff script and then fathom to generate the export.ann and export.dna files, we will use 
a script that converts the zff files to GenBank format to train Augustus using the zff2augustus_gbk.pl script available at:  
https://github.com/hyphaltip/genome-scripts/blob/master/gene_prediction/zff2augustus_gbk.pl.  
  
#### 6-A. Set up the training directories:
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$maker_rnd2_path[k], "rnd2_Augustus_Training", sep = ""), showWarnings = TRUE, recursive = FALSE)}

metadataAssemblies$rnd2_AugTrain <- paste(metadataAssemblies$maker_rnd2_path, "rnd2_Augustus_Training/", sep = "")
```

#### 6-B. Convert the zff files to GenBank format to prepare for Augustus training:
```{r}
prefix <- "round2_Augustus_zff2gbk_Fustric"

cmd    <- with(metadataAssemblies, 
               paste("cd ", rnd2_AugTrain, " && ", zff2augGbkPath, 
                     " ", paste(snapRnd1MakerRnd2Path, "export.ann", sep = ""),
                     " ", paste(snapRnd1MakerRnd2Path, "export.dna", sep = ""),
                     " > ", SppAbbrv, ".train.gb",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

metadataAssemblies$rnd2AugTrainGB     <- paste(metadataAssemblies$SppAbbrv, ".train.gb", sep = "")
metadataAssemblies$rnd2AugTrainGBpath <- paste(metadataAssemblies$rnd2_AugTrain, metadataAssemblies$rnd2AugTrainGB, sep = "")
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Not sure of the following steps, so they are test and need to be reviewed. They follow the steps
outlined in https://vcru.wisc.edu/simonlab/bioinformatics/programs/augustus/docs/tutorial2015/training.html.  
  
#### 6-C Randomly split the set of annotated sequences in a training and a test set:  
In order for the test accuracy to be statistically meaningful the test set should also be large enough (100-200 genes). 
We need to split the set of gene structures really randomly! Do not just take the first and the last part of the file 
as then the test set is unlikely to be representative. We can achieve this with the randomSplit.pl script that comes
with the Augustus package. 
```{r}
prefix <- "round2_Aug_randomSplit_Fustric"

cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", rnd2_AugTrain, " && ", "randomSplit.pl ", rnd2AugTrainGB, " 100 ",
                     " && conda deactivate ",
                     sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 6-D. Set up the meta parameters file and create the species training files.  
**Make a meta parameters file for our species.**  
We call parameters like the size of the window of the splice site models and the order of the Markov model 
meta parameters, in contrast to parameters like the distribution of splice site patterns, the k-mer probabilities 
of coding and noncoding regions. There are a few dozen meta parameters but many thousands of parameters. The meta 
parameters determine how the parameters are calculated. 
  
**Create the files for training our species from a template:**  
new_species.pl uses the environment variable AUGUSTUS_CONFIG_PATH to determine the directory in which 
AUGUSTUS stores the species parameters.
```{r}
prefix <- "round2_AugTrainingFiles_Fustric"
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", rnd2_AugTrain, " && new_species.pl ",
                     " --species=", ScientificName, "_AugRnd2",
                     " --AUGUSTUS_CONFIG_PATH=/isilon/cfia-ottawa-fallowfield/users/girouxeml/prog/augustus/config",
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 6-E. Perform an inital training run.  
We will use the training set of gene structures, genes.train.gb, to train Augustus for our species. Etraining will create/update
parameter files for exon, intron, and intergenic regions in the $AUGUSTUS_CONFIG_PATH/species/ourSpecies.
```{r}
prefix <- "round2_AugEtraining_Fustric"
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", rnd2_AugTrain, " && etraining ",
                     " --species=", ScientificName, "_AugRnd2 ", rnd2AugTrainGB,
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```
[CFIA-ACIA+girouxeml@biocomp-0-5 round2_AugEtraining_Fustric]$ cat round2_AugEtraining_Fustric1.sub.o284796
# Read in 7296 genbank sequences.
Quantiles of the GC contents in the training set:
0%      0.27
5%      0.444   10%     0.456
15%     0.462   20%     0.468
25%     0.472   30%     0.476
35%     0.479   40%     0.482
45%     0.485   50%     0.488
55%     0.491   60%     0.494
65%     0.497   70%     0.5
75%     0.503   80%     0.506
85%     0.51    90%     0.516
95%     0.523   100%    0.566
HMM-training the parameters...
i= 0 bc= (0.237, 0.263, 0.263, 0.237)
 ** building model for exons *EXON*
start codon frequencies: ATG(7296)
# admissible start codons and their probabilities: ATG(1), CTG(0), TTG(0)
 number of bases in the reading frames: 3793062 3800299 3800306
--- frame = 0 ---    minPatSum = 233
--- frame = 1 ---    minPatSum = 233
--- frame = 2 ---    minPatSum = 233
--- initial frame = 0 ---    minPatSum = 233
--- initial frame = 1 ---    minPatSum = 233
--- initial frame = 2 ---    minPatSum = 233
--- internal exon terminal frame = 0 ---    minPatSum = 233
--- internal exon terminal frame = 1 ---    minPatSum = 233
--- internal exon terminal frame = 2 ---    minPatSum = 233
single, initial, internal, terminal mean exon lengths :
n.a.    312     336     573
single exon : 0
initial exon 0 : 2870
initial exon 1 : 2483
initial exon 2 : 1943
internal exon 0 : 3807
internal exon 1 : 2831
internal exon 2 : 2693
terminal exon : 7296
Frequency of stop codons:
tag: 2373 (0.325)
taa: 2798 (0.383)
tga: 2125 (0.291)
end *EXON*
Storing parameters to file...
Writing exon model parameters [1] to file /home/CFIA-ACIA/girouxeml/prog/miniconda/envs/makerenv/config/species/Fusarium_tricinctum_AugRnd2/Fusarium_tricinctum_AugRnd2_exon_probs.pbl.

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 6-F. Test the prediction accuracy of the newly-created Augustus training files.  
We can test the prediction accuracy of our new Augustus training files on the test set we generated when we used
randomSplit.pl.
```{r}
prefix <- "round2_AugTrain_InitTest_Fustric"
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", rnd2_AugTrain, " && augustus ", " --species=", ScientificName, "_AugRnd2 ",
                     rnd2AugTrainGB, ".test  2>&1 | tee firsttest.out",
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```
**Checking the Evaluation Results.**  
You can see the evaluation summary in the firsttest.out file by typing the following in the command line:  
  
**_$ grep -A 22 Evaluation firsttest.out_**  
  
You should see an output like the following:  
  
*******      Evaluation of gene prediction     *******

---------------------------------------------\
                 | sensitivity | specificity |
---------------------------------------------|
nucleotide level |       0.969 |       0.956 |
---------------------------------------------/

----------------------------------------------------------------------------------------------------------\
           |  #pred |  #anno |      |    FP = false pos. |    FN = false neg. |             |             |
           | total/ | total/ |   TP |--------------------|--------------------| sensitivity | specificity |
           | unique | unique |      | part | ovlp | wrng | part | ovlp | wrng |             |             |
----------------------------------------------------------------------------------------------------------|
           |        |        |      |                 58 |                 45 |             |             |
exon level |    344 |    331 |  286 | ------------------ | ------------------ |       0.864 |       0.831 |
           |    344 |    331 |      |   33 |    2 |   23 |   30 |    0 |   15 |             |             |
----------------------------------------------------------------------------------------------------------/

----------------------------------------------------------------------------\
transcript | #pred | #anno |   TP |   FP |   FN | sensitivity | specificity |
----------------------------------------------------------------------------|
gene level |   109 |   100 |   71 |   38 |   29 |        0.71 |       0.651 |
----------------------------------------------------------------------------/

#### 6-G. Optimize the Augustus prediction accuracy.  
We can optimize the Augustus prediction accuracy by running the optimize_augustus.pl script that comes with the Augustus
package. This script adjusts the meta parameters in the *_parameters.cfg file. Note that this script will alternatingly
use augustus and etraining programs, and may take a long time to complete. It can increase the prediction accuracy by
a few percentage points.  
```{r}
prefix <- "round2_AugTrain_Optimize_cpus12_Fustric"
node   <- 12
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", rnd2_AugTrain, " && optimize_augustus.pl", " --cpus=", node, 
                     " --species=", ScientificName, "_AugRnd2 ", rnd2AugTrainGB, ".train",
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```
If you want to see how the improvements from training are coming along, you can use awk onthe file:  
[CFIA-ACIA+girouxeml@biocomp-0-2 round2_AugTrain_Optimize_cpus12_Fustric]$ awk '/starting accuracy:|found improvement/ { print }' round2_AugTrain_Optimize_cpus12_Fustric1.sub.o284798
starting accuracy: 0.9630, 0.9427, 0.8343, 0.8044, 0.6659, 0.5916, -1.00, -1.00, starting target: 0.8419
found improvement: 0.9654, 0.9416, 0.8415, 0.8051, 0.6709, 0.5955, -1.00, -1.00, optimal target: 0.8446
found improvement: 0.9644, 0.9419, 0.8410, 0.8065, 0.6717, 0.5969, -1.00, -1.00, optimal target: 0.8449
found improvement: 0.9650, 0.9419, 0.8421, 0.8059, 0.6724, 0.5968, -1.00, -1.00, optimal target: 0.8451
found improvement: 0.9657, 0.9419, 0.8503, 0.8121, 0.6908, 0.6113, -1.00, -1.00, optimal target: 0.8514
found improvement: 0.9659, 0.9422, 0.8504, 0.8125, 0.6912, 0.6119, -1.00, -1.00, optimal target: 0.8517
found improvement: 0.9664, 0.9420, 0.8518, 0.8131, 0.6939, 0.6134, -1.00, -1.00, optimal target: 0.8525
found improvement: 0.9664, 0.9421, 0.8518, 0.8132, 0.6940, 0.6136, -1.00, -1.00, optimal target: 0.8526
found improvement: 0.9662, 0.9421, 0.8520, 0.8134, 0.6940, 0.6138, -1.00, -1.00, optimal target: 0.8527
found improvement: 0.9664, 0.9421, 0.8522, 0.8135, 0.6944, 0.6140, -1.00, -1.00, optimal target: 0.8528
found improvement: 0.9665, 0.9421, 0.8524, 0.8136, 0.6943, 0.6139, -1.00, -1.00, optimal target: 0.8529
found improvement: 0.9666, 0.9421, 0.8528, 0.8141, 0.6947, 0.6144, -1.00, -1.00, optimal target: 0.8532
found improvement: 0.9666, 0.9421, 0.8531, 0.8141, 0.6950, 0.6145, -1.00, -1.00, optimal target: 0.8533
found improvement: 0.9667, 0.9425, 0.8529, 0.8141, 0.6952, 0.6156, -1.00, -1.00, optimal target: 0.8534
found improvement: 0.9667, 0.9426, 0.8534, 0.8159, 0.6961, 0.6162, -1.00, -1.00, optimal target: 0.8540
found improvement: 0.9664, 0.9426, 0.8544, 0.8164, 0.6985, 0.6191, -1.00, -1.00, optimal target: 0.8548
found improvement: 0.9664, 0.9427, 0.8545, 0.8165, 0.6995, 0.6196, -1.00, -1.00, optimal target: 0.8550
found improvement: 0.9665, 0.9427, 0.8547, 0.8165, 0.7002, 0.6202, -1.00, -1.00, optimal target: 0.8552
found improvement: 0.9665, 0.9426, 0.8547, 0.8167, 0.7002, 0.6204, -1.00, -1.00, optimal target: 0.8553
found improvement: 0.9667, 0.9427, 0.8553, 0.8166, 0.7005, 0.6205, -1.00, -1.00, optimal target: 0.8555
found improvement: 0.9661, 0.9429, 0.8554, 0.8171, 0.7014, 0.6213, -1.00, -1.00, optimal target: 0.8557
found improvement: 0.9665, 0.9427, 0.8559, 0.8172, 0.7025, 0.6222, -1.00, -1.00, optimal target: 0.8560
found improvement: 0.9665, 0.9427, 0.8559, 0.8174, 0.7024, 0.6221, -1.00, -1.00, optimal target: 0.8561
found improvement: 0.9665, 0.9427, 0.8560, 0.8175, 0.7027, 0.6225, -1.00, -1.00, optimal target: 0.8562
found improvement: 0.9665, 0.9427, 0.8560, 0.8177, 0.7031, 0.6228, -1.00, -1.00, optimal target: 0.8563
found improvement: 0.9715, 0.9414, 0.8606, 0.8187, 0.7064, 0.6216, -1.00, -1.00, optimal target: 0.8582
found improvement: 0.9716, 0.9412, 0.8610, 0.8186, 0.7071, 0.6228, -1.00, -1.00, optimal target: 0.8584

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 6-H. Re-train Augustus using the new meta parameters it has set    
Once the optimize_augustus.pl has finished (or been interrupted), we need to re-train Augustus with the newly
optimized meta parameters. Note - after running the following chunk, it was recommended to change the
parameters "I recommend to set /Constant/gc_range_min to 0.44 and /Constant/gc_range_max to 0.52" - which I did with gedit:   
$ ~/prog/miniconda/envs/makerenv/config/species/Fusarium_tricinctum_AugRnd2/Fusarium_tricinctum_AugRnd2_parameters.cfg   
```{r}
prefix <- "round2_AugEtraining_2_Fustric"
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", rnd2_AugTrain, " && etraining ",
                     " --species=", ScientificName, "_AugRnd2 ", rnd2AugTrainGB,
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 6-I. Test the prediction accuracy of the newly-created Augustus training files.    
We can test the prediction accuracy of our new Augustus training files on the test set we generated when we used
randomSplit.pl.
```{r}
prefix <- "round2_AugTrain_2ndTest_Fustric"
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", rnd2_AugTrain, " && augustus  --species=", ScientificName, "_AugRnd2 ",
                     rnd2AugTrainGB, ".test  2>&1 | tee secondtest.out",
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

*******      Evaluation of gene prediction     *******

---------------------------------------------\
                 | sensitivity | specificity |
---------------------------------------------|
nucleotide level |       0.969 |       0.956 |
---------------------------------------------/

----------------------------------------------------------------------------------------------------------\
           |  #pred |  #anno |      |    FP = false pos. |    FN = false neg. |             |             |
           | total/ | total/ |   TP |--------------------|--------------------| sensitivity | specificity |
           | unique | unique |      | part | ovlp | wrng | part | ovlp | wrng |             |             |
----------------------------------------------------------------------------------------------------------|
           |        |        |      |                 63 |                 45 |             |             |
exon level |    349 |    331 |  286 | ------------------ | ------------------ |       0.864 |       0.819 |
           |    349 |    331 |      |   35 |    3 |   25 |   31 |    1 |   13 |             |             |
----------------------------------------------------------------------------------------------------------/

----------------------------------------------------------------------------\
transcript | #pred | #anno |   TP |   FP |   FN | sensitivity | specificity |
----------------------------------------------------------------------------|
gene level |   110 |   100 |   72 |   38 |   28 |        0.72 |       0.655 |
----------------------------------------------------------------------------/

------------------------------------------------------------------------\
            UTR | total pred | CDS bnd. corr. |   meanDiff | medianDiff |
------------------------------------------------------------------------|
            TSS |          3 |              0 |         -1 |         -1 |
            TTS |          2 |              0 |         -1 |         -1 |
------------------------------------------------------------------------|
            UTR | uniq. pred |    unique anno |      sens. |      spec. |
------------------------------------------------------------------------|
                |  true positive = 1 bound. exact, 1 bound. <= 20bp off |
 UTR exon level |          0 |              0 |       -nan |       -nan |
------------------------------------------------------------------------|
 UTR base level |          0 |              0 |       -nan |       -nan |
------------------------------------------------------------------------/
nucUTP= 0 nucUFP=0 nucUFPinside= 0 nucUFN=0

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

### 7. Run MAKER for a third round.  
We now will run MAKER a third time. We will be able to use the re-trained Augustus models, as well as the other SNAP and
Genemark-ES models we previously generated, tying all our optimizations together. The idea is to run MAKER until the results
do not vary much between successive rounds of MAKER, signalling that we have optimized the predictions and output for our data.  

#### 7-A. Recycle the mapping and empirical evidence obtained from the second round with MAKER:  
Note that we are not generating a protein2genome.gff3 set since we set protein2genome to 0 in MAKER round 2, however, we do have
updated evidence for repeats, so will generate that file.  
```{r}
# Repeat alignments:
prefix <- "maker_rnd3_repeat_evidence_Fustric"
cmd    <- with(metadataAssemblies,
               paste("cd ", makerRnd2_OutputPath, " && ",
                     paste(" awk '{ if ", '($2 ~ "repeat") print $0 }', "'", sep = ""),
                     " ", paste(SppAbbrv, "_rnd2.all.maker.noseq.gff3", sep = ""), 
                     " > ", paste(SppAbbrv, "_rnd2.all.maker.repeats.gff3", sep = ""),
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 7-B. Create the directory to contain MAKER round 3 output:  
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[k], "maker_round_3", sep = ""), 
               showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$maker_rnd3_path <- paste(metadataAssemblies$RepeatDBPath, "maker_round_3/", sep = "")
```

#### 7-C. Generate the MAKER round 3 control files:
```{r}
prefix <- "maker_rnd3_CTL_Fustric"
cmd    <- with(metadataAssemblies, 
               paste("conda activate makerenv && ",
                     "cd ", maker_rnd3_path, " && maker  -CTL ",
                     " && conda deactivate ",
                     sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
metadataAssemblies$makerRnd3_CTL_optsPath <- paste(metadataAssemblies$maker_rnd3_path, "maker_opts.ctl", sep = "")
metadataAssemblies$makerRnd3_CTL_exePath  <- paste(metadataAssemblies$maker_rnd3_path, "maker_exe.ctl", sep = "")
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 7-D. Edit the MAKER round 3 control files.  
Modify the control file, removing the FASTA sequences files to map and replacing them with the GFFs (est_gff, 
protein_gff, and rm_gff, respectively. We can also specify the path to the Genemark-es .mod, SNAP HMM and the 
species name for Augustus that we re-trained, so that these gene prediction programs are run. We will also switch est2genome and 
protein2genome to 0 so that gene predictions are based on the Augustus and SNAP gene models.
```{r}
cmd <- with(metadataAssemblies,
            paste("/opt/rocks/bin/replace ",
                  " altest= ", paste("altest=", cdnaFuspoaPath, sep = ""),
                  " altest_gff= ", paste("altest_gff=", gff3FuspoaPath, sep = ""),
                  " protein_gff= ", paste("protein_gff=", makerRnd1_OutputPath, SppAbbrv, "_rnd1.all.maker.protein2genome.gff3", sep = ""),
                  " model_org=all model_org=# ",
                  " repeat_protein= repeat_protein=# ",
                  " rm_gff= ", paste("rm_gff=", makerRnd2_OutputPath, SppAbbrv, "_rnd2.all.maker.repeats.gff3", sep = ""),
                  " snaphmm= ", paste("snaphmm=", snapRnd1MakerRnd2Path, SppAbbrv, "_rnd2.zff.hmm", sep = ""),
                  " gmhmm= ", paste("gmhmm=", genemarkEsRnd1Path, "output/", SppAbbrv, "_GMES.mod", sep = ""),
                  " augustus_species= ", paste("augustus_species=", ScientificName, "_AugRnd2", " ", sep = ""),
                  " trna=0 trna=1 ",
                  " max_dna_len=100000 max_dna_len=300000 ",
                  " -- ", makerRnd3_CTL_optsPath,
                  sep = ""))
sapply(cmd, function(x) system(x))

# I also modified the maker_exe.ctl with the local path to tRNAscan: If this isn't corrected, and the 
# conda makerenv trnascan is used, the analysis will fail.

pattern    <- paste("tRNAscan-SE=", pathtRNA_scan, sep = "")
newPattern <- paste("tRNAscan-SE=", pathtRNA_scan, sep = "")
cmd        <- with(metadataAssemblies, paste("/opt/rocks/bin/replace ", pattern, " ", newPattern,
                                             " -- ", makerRnd3_CTL_exePath, sep = ""))
sapply(cmd, function(x) system(x))

# Edit Genome paths:
pattern    <- "genome= "
newPattern <- paste("genome=", metadataAssemblies$FixedNewAssemName, sep = "")
for(i in 1:nrow(metadataAssemblies)){
    makerOptsGenRepl(metadataAssemblies$makerRnd3_CTL_optsPath[i], pattern, newPattern[i])}
```

#### 7-E. Run MAKER round 3.
**Then we can run MAKER, substituting this new control file, and summarize the output, as we did before.**
```{r}
prefix <- "maker_round_3_run_Fustric"
node   <- 12
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", maker_rnd3_path, " && maker ", 
                     " -base ", paste(SppAbbrv, "_rnd3", sep = ""),
                     " ", paste(maker_rnd3_path, "maker_opts.ctl", sep = ""),
                     " ", paste(maker_rnd3_path, "maker_bopts.ctl", sep = ""),
                     " ", paste(maker_rnd3_path, "maker_exe.ctl", sep = ""),
                     " -cpus ", node,
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

#### 7-F. Assemble the MAKER round 3 GFF and FASTA outputs.  
**We conclude by assembling together the GFF and FASTA outputs.**
```{r}
metadataAssemblies$makerRnd3_OutputPath <- paste(metadataAssemblies$maker_rnd3_path, 
                                                 metadataAssemblies$SppAbbrv, "_rnd3.maker.output/", sep = "")

# GFF with the sequences:
prefix  <- "round3_mergeGFF3_maker_all_Fustric"
cmd     <- with(metadataAssemblies,
                paste("conda activate makerenv && ",
                      "cd ", makerRnd3_OutputPath, " && gff3_merge -s -d ",
                      paste(makerRnd3_OutputPath, SppAbbrv, "_rnd3_master_datastore_index.log", sep = ""),
                      " > ", paste(makerRnd3_OutputPath, SppAbbrv, "_rnd3.all.maker.gff3", sep = ""),
                      " && conda deactivate ",
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

# Collect all the fastas:
prefix2 <- "round3_fastaMerge_maker_Fustric"
cmd     <- with(metadataAssemblies,
                paste("conda activate makerenv && ",
                      "cd ", makerRnd3_OutputPath, " && fasta_merge -d ",
                      paste(makerRnd3_OutputPath, SppAbbrv, "_rnd3_master_datastore_index.log", sep = ""),
                      " && conda deactivate ",
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix2, suffix)

# GFF without the sequences:
prefix3 <- "round3_mergeGFF3_maker_all_noseq_Fustric"
cmd     <- with(metadataAssemblies,
                paste("conda activate makerenv && ",
                      "cd ", makerRnd3_OutputPath, " && gff3_merge -n -s -d ",
                      paste(makerRnd3_OutputPath, SppAbbrv, "_rnd3_master_datastore_index.log", sep = ""),
                      " > ", paste(makerRnd3_OutputPath, SppAbbrv, "_rnd3.all.maker.noseq.gff3", sep = ""),
                      " && conda deactivate ",
                      sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix3, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2); RemoveQsubTempFiles(sharedPathAn, prefix3)
```

**Export confident gene models from round 3 of MAKER using maker2zff:**  
For determining which genes are High Confidence for Retraining:
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$maker_rnd3_path[k], "snap", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$snapMakerRnd3Path <- paste(metadataAssemblies$maker_rnd3_path, "snap/", sep = "")

for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$snapMakerRnd3Path[k], "round1", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$snapRnd1MakerRnd3Path <- paste(metadataAssemblies$snapMakerRnd3Path, "round1/", sep = "")

prefix <- "round3_snapMaker2zff_Fustric"
cmd    <- with(metadataAssemblies,
               paste("conda activate makerenv && ",
                     "cd ", snapRnd1MakerRnd3Path, " && maker2zff  -x 0.25 -l 50 -c 0 -e 0 -o 0 -d ", 
                     paste(makerRnd3_OutputPath, SppAbbrv, "_rnd3_master_datastore_index.log", sep = ""),
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Take a quick look at number of sequences in the genome.dna file:**
```{r}
cmd <- with(metadataAssemblies, paste("cd ", snapRnd1MakerRnd3Path, " && ", 'grep ">" genome.dna | wc -l ', sep = ""))
sapply(cmd, function(x) system(x))

cmd <- with(metadataAssemblies, file.rename(paste(snapRnd1MakerRnd3Path, "genome.dna", sep = ""),
                                            paste(snapRnd1MakerRnd3Path, SppAbbrv, "_rnd3.zff.dna", sep = "")))
sapply(cmd, function(x) system(x))

cmd <- with(metadataAssemblies, file.rename(paste(snapRnd1MakerRnd3Path, "genome.ann", sep = ""),
                                            paste(snapRnd1MakerRnd3Path, SppAbbrv, "_rnd3.zff.ann", sep = "")))
sapply(cmd, function(x) system(x))
```

**Gather some stats and validate:**
```{r}
prefix  <- "round3_fathom_stats_Fustric"
cmd     <- with(metadataAssemblies, 
                paste("conda activate makerenv && ",
                      "cd ", snapRnd1MakerRnd3Path, " && fathom  ", 
                      paste(SppAbbrv, "_rnd3.zff.ann", sep = ""), " ", paste(SppAbbrv, "_rnd3.zff.dna", sep = ""),
                      " -gene-stats > gene-stats.log 2>&1 ",
                      " && conda deactivate ",
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)

prefix2 <- "round3_fathom_validate_Fustric"
cmd2    <- with(metadataAssemblies, 
                paste("conda activate makerenv && ",
                      "cd ", snapRnd1MakerRnd3Path, " && fathom ", 
                      paste(SppAbbrv, "_rnd3.zff.ann", sep = ""), " ", paste(SppAbbrv, "_rnd3.zff.dna", sep = ""),
                      " -validate > validate.log 2>&1 ",
                      " && conda deactivate ",
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix); RemoveQsubTempFiles(sharedPathAn, prefix2)
```

**Collect the training sequences and annotations, plus 1000 surrounding bp for training**
```{r}
prefix <- "round3_fathom_categorize_Fustric"
cmd    <- with(metadataAssemblies, 
               paste("conda activate makerenv && ",
                     "cd ", snapRnd1MakerRnd3Path, " && fathom ", 
                     paste(SppAbbrv, "_rnd3.zff.ann", sep = ""), " ", paste(SppAbbrv, "_rnd3.zff.dna", sep = ""),
                     " -categorize 1000 > categorize.log 2>&1 ",
                     " && conda deactivate ",
                     sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Wait for the categorize run to finish before doing the export, which is next.**
```{r}
prefix2 <- "round3_fathom_export_Fustric"
cmd2    <- with(metadataAssemblies,
                paste("conda activate makerenv && ",
                      "cd ", snapRnd1MakerRnd3Path, 
                      " && fathom uni.ann uni.dna -export 1000 -plus > uni-plus.log 2>&1 ",
                      " && conda deactivate ",
                      sep = ""))
suffix  <- ".sub"; cat(bashDirections); MakeQsubs(cmd2, prefix2, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix2)
```

### We've now completed the 3rd round of MAKER.   
#### 7-G. Evaluate the gene models.  
We've now completed the 3rd round of MAKER and should evaluate the gene models we have so far:  

**i. Count the number of gene models and the gene lengths:**    
I manually recorded the output values to a summary BUSCO excel file. Welcome to ideas on how to capture the output.
```{r}
cmd <- with(metadataAssemblies, 
            paste("cd ", makerRnd3_OutputPath, " && ", " cat ", SppAbbrv, "_rnd3.all.maker.gff3 ", 
                  " | awk '{ if ($3 ==", '"gene") print $0}', 
                  "' | awk '{ sum += ($5 - $4) } END { print NR, sum / NR }'", sep = ""))

sapply(cmd, function(x) system(x))
```
13764 1649.97
cd /isilon/cfia-ottawa-fallowfield/users/girouxeml/PIRL_working_directory/Fusarium_Pramorum_species_GenomeAn_IonTorrent_2018/Fusarium_tricinctum/Fustric_RepeatDB/maker_round_3/Fustric_rnd3.maker.output/ &&  cat Fustric_rnd3.all.maker.gff3  | awk '{ if ($3 =="gene") print $0}' | awk '{ sum += ($5 - $4) } END { print NR, sum / NR }' 
                                                                                                                                                                                                                                                                                                                                           0 
**ii. Visualize the AED distribution:**   
AED ranges from 0 to 1 and quantifies the confidence in a gene model based on empirical evidence. Basically, 
the lower the AED, the better a gene model is likely to be. Ideally, 95% or more of the gene models will have an 
AED of 0.5 or better in the case of good assemblies. You can use this AED_cdf_generator.pl script to help with this.  
I have just recorded the fraction at 0.5 into my summary excel table. Open to ideas on how to capture it better. 
```{r}
cmd <- with(metadataAssemblies, 
            paste("cd ", makerRnd3_OutputPath, " && ", " perl ", aedCDFgenePath, 
                  " -b 0.025 ", SppAbbrv, "_rnd3.all.maker.gff3", sep = "")) 
sapply(cmd, function(x) system(x))
```
AED	Fustric_rnd3.all.maker.gff3
0.000	0.397
0.025	0.632
0.050	0.748
0.075	0.783
0.100	0.812
0.125	0.826
0.150	0.845
0.175	0.858
0.200	0.874
0.225	0.883
0.250	0.900
0.275	0.911
0.300	0.926
0.325	0.937
0.350	0.950
0.375	0.959
0.400	0.970
0.425	0.976
0.450	0.983
0.475	0.986
0.500	0.990
0.525	0.991
0.550	0.992
0.575	0.993
0.600	0.995
0.625	0.995
0.650	0.996
0.675	0.997
0.700	0.997
0.725	0.998
0.750	0.999
0.775	0.999
0.800	0.999
0.825	0.999
0.850	1.000
0.875	1.000
0.900	1.000
0.925	1.000
0.950	1.000
0.975	1.000
cd /isilon/cfia-ottawa-fallowfield/users/girouxeml/PIRL_working_directory/Fusarium_Pramorum_species_GenomeAn_IonTorrent_2018/Fusarium_tricinctum/Fustric_RepeatDB/maker_round_3/Fustric_rnd3.maker.output/ &&  perl /home/CFIA-ACIA/girouxeml/prog/scripts_pl/AED_cdf_generator.pl -b 0.025 Fustric_rnd3.all.maker.gff3 1.000	1.000


**iii. Run BUSCO using the Augustus species HMM to look at the results:**  
We can run BUSCO using the Augustus species HMM we obtained, without having to train Augustus. Only include the 
transcript sequences, and not the 1000 bp on each side, and be sure to take the best (i.e., longest) transcript for
each gene so we aren't artificially seeding duplicates. We could also run it on the best protein sequence per gene
instead.  
```{r}
prefix <- "round3_eval_with_Busco_Fustric"
node   <- 1
cmd    <- with(metadataAssemblies,
               paste("cd ", makerRnd3_OutputPath, " && ",
                     "python ", buscoPath, " -i ", paste(SppAbbrv, "_rnd3.all.maker.transcripts.fasta", sep = ""),
                     " -o annotation_eval ", " -c ", node,
                     " -l ", buscoSorDataSet, " -m transcriptome -sp ", ScientificName, "_AugRnd2", " -z --augustus_parameters='--progress=true'",
                     " --restart  --force ", sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
# Note, if the run fails, restart the run with the following "--restart", it will continue from where it left off. However, check the 
# config.ini file in ~/prog/busco/config/ to make sure that restart is set to True, and that you adjust to the correct cpus number.
```
INFO    Results:
INFO    C:97.9%[S:97.6%,D:0.3%],F:1.3%,M:0.8%,n:3725
INFO    3647 Complete BUSCOs (C)
INFO    3637 Complete and single-copy BUSCOs (S)
INFO    10 Complete and duplicated BUSCOs (D)
INFO    48 Fragmented BUSCOs (F)
INFO    30 Missing BUSCOs (M)
INFO    3725 Total BUSCO groups searched

D. Visualise the gene models from Augustus, SNAP, and MAKER, also Genemark-es using a genome browser such as JBrowse or Geneious.  
It is best to use 5-10 longer, gene dense scaffolds and visually inspect them. When SNAP and Augustus are well trained, their models 
should overlap pretty closely with the final MAKER models. Moreover, there will be spurious hits from SNAP and Augustus, but they are 
usually short, 1-2 exon annotations and don't have empirical support. You'll get a sense of a good annotation with some experience. 
Also, it is possible SNAP won't produce good results, depending on your organism, which the MAKER folks have pointed out in the 
past (Augustus usually does pretty well).
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[k], "jbrowse", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$jbrowse <- paste(metadataAssemblies$RepeatDBPath, "jbrowse/", sep = "")

prefix <-"end_Jbrowse_prepRefSeqs"
cmd    <- with(metadataAssemblies,
               paste("cd ", jbrowse, " && ", jbrowsePrepRefsSeqsPath, 
                     " --gff ", paste(makerRnd3_OutputPath, SppAbbrv, "_rnd3.all.maker_jb.gff3", sep = "")))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
# Have yet to determine quickest way to do this on our cluster.
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Compress Maker round 1 and 2 output directories**  
We will be using files from MAKER round 3 for processing from now on, and so should compress these 
directories to conserve space. If necessary, we can untar them and refer back to the files.
```{r}
prefix <- "tar_zcvf_maker_round_2_3_Fustric"
cmd    <- with(metadataAssemblies, 
               paste("tar -zcvf ", paste(RepeatDBPath, "maker_round_1.tar.gz ", sep = ""), 
                     paste(RepeatDBPath, "maker_round_1", sep = ""), "; ",
                     "tar -zcvf ", paste(RepeatDBPath, "maker_round_2.tar.gz ", sep = ""),
                     paste(RepeatDBPath, "maker_round_2", sep = ""), "; ",
                     "rm -rf ", paste(RepeatDBPath, "maker_round_1", sep = ""), "; ", 
                     "rm -rf ", paste(RepeatDBPath, "maker_round_2", sep = ""), sep = ""))
suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix, node)
```
**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

**Prepare for post-MAKER processing***
```{r}
for(k in 1:nrow(metadataAssemblies)){
    dir.create(paste(metadataAssemblies$RepeatDBPath[k], "postMakerProcessing", sep = ""), showWarnings = TRUE, recursive = FALSE)}
metadataAssemblies$postMaker <- paste(metadataAssemblies$RepeatDBPath, "postMakerProcessing/", sep = "")
```

Copy the GFF3 files output from Maker round 3 to the new directory just created.
```{r}
cmd <- with(metadataAssemblies, 
            paste("cp ", paste(metadataAssemblies$makerRnd3_OutputPath, metadataAssemblies$SppAbbrv, "_rnd3.all.maker.gff3", sep = ""), 
                  " ", postMaker, sep = ""))

sapply(cmd, function(x) system(x))
```
Copy the trasncripts, proteins, and tRNA fasta files that were generated from the last round of MAKER
to the new post MAKER directory, then compress files in the last MAKER run directory.
```{r}
prefix <- "postMaker_cp_gz_fasta_Fustric"
cmd    <- with(metadataAssemblies, 
               paste("cp ", paste(metadataAssemblies$makerRnd3_OutputPath, metadataAssemblies$SppAbbrv,
                                  "_rnd3.all.maker.transcripts.fasta", sep = ""), 
                     " ", paste(postMaker, "/; ", sep = ""), 
                     "cp ", paste(metadataAssemblies$makerRnd3_OutputPath, metadataAssemblies$SppAbbrv,
                                  "_rnd3.all.maker.proteins.fasta", sep = ""), 
                     " ", paste(postMaker,"/; ", sep = ""),
                     "cp ", paste(metadataAssemblies$makerRnd3_OutputPath, metadataAssemblies$SppAbbrv,
                                  "_rnd3.all.maker.trnascan.transcripts.fasta", sep = ""), 
                     " ", paste(postMaker, "/;", sep = ""),
                     "gzip ", paste(metadataAssemblies$makerRnd3_OutputPath, "*", sep = ""), sep = ""))

suffix <- ".sub"; cat(bashDirections); MakeQsubs(cmd, prefix, suffix)
```

**To remove the output files after you are done:**
```{r}
RemoveQsubTempFiles(sharedPathAn, prefix)
```

Record the JBrowse-fixed gff3 files in the metadata table: *Note - not using jbrowse, but adding name to 
metadata table so preserve script flow from hereon.
```{r}
metadataAssemblies$fixed_GFF3_jb <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, "_rnd3.all.maker.gff3", sep = "")
```

### Generation of NCBI Submission Files with GAG - to validate and fix gff errors:    
We can generate our genome submission files using the Genome Annotation Generator (GAG) prior to running BLAST and InterProScan.
What we focus on from this are the validation and error files that are generated that will tell us all the gene features that are invalid, such as protein with internal stop detected, or incomplete gene or missing stop codn, etc., and we go through these and fix all the errors, re-run GAG iteratively until no more errors are present. Once all the errors are solved, then we use InterProscan and Blast and whatever else for annotation, and so dont waste our time annotating invalid features. 

$ mkdir gag; cd gag  
$  python ~/prog/gag/genomeannotation-GAG-40ea515/gag.py -f ../../../../../References/Fusarium_tricinctum_GCA_900382705.2/Fusarium_tricinctum_GCA_900382705.2_FTRI.INRA104.GCA2018.2_genomic.fna -g ../Fustric_rnd3.all.maker.gff3 --fix_start_stop -ris 10 -rcs 75 -o gag_pass1_fix_start_stop_short_introns    

$ mv genome.tbl Fustric.tbl     
$ mv genome.fasta Fustric.fsa   
$ mkdir tbl2asn  
$ cp ../../../../../template.sbt tbl2asn/Fustric_template.sbt
$ cp Fustric* tbl2asn/  
$ cd tbl2asn   
$ ~/prog/linux64.tbl2asn -j "[organism=Fusarium_tricinctum] [strain=ABC 123]" -t Fustric_template.sbt -p. -M n -Z discrep -a r1k -V b -X E  
$ cat errorsummary.val
 26524 ERROR:   SEQ_DESCR.DBLinkProblem
   319 ERROR:   SEQ_FEAT.MissingTrnaAA
   546 WARNING: SEQ_FEAT.NotSpliceConsensusAcceptor
   154 WARNING: SEQ_FEAT.NotSpliceConsensusDonor
  1540 WARNING: SEQ_FEAT.PartialProblem
   146 WARNING: SEQ_FEAT.ShortExon
    86 INFO:    SEQ_FEAT.PartialProblem
     2 INFO:    SEQ_FEAT.RareSpliceConsensusDonor

Fix the errors captured in this document. I didn't fix the tRNA errors, it's not 
necessary to fix them yet since we are not yet going for NCBI submission.
Copy the gff and fasta files from the final GAG edits folder, into the postMakerProcessing directory, with the species 
abbreviation suffix.
```{r}
previousGagDir <- "gag_pass1_fix_start_stop_short_introns/"
cmd <- with(metadataAssemblies,
            paste("cp ", paste(postMaker, "gag/", previousGagDir, "genome.edit.gff", sep = ""),
                  " ", paste(postMaker, SppAbbrv, "_gag_edited.gff", sep = ""),  " ;",
                  "cp ", paste(postMaker, "gag/", previousGagDir, SppAbbrv, ".fsa", sep = ""),
                  " ", paste(postMaker, SppAbbrv, "_gag_edited.fsa", sep = ""), " ;",
                  "cp ", paste(postMaker, "gag/", previousGagDir, SppAbbrv, ".edit.tbl", sep = ""),
                  " ", paste(postMaker, SppAbbrv, ".tbl", sep = ""), " ;",
                  "cp ", paste(postMaker, "gag/", previousGagDir, "genome.proteins.fasta", sep = ""),
                  " ", paste(postMaker, SppAbbrv, ".proteins_gag_edited.fasta", sep = ""), " ;",
                  "cp ", paste(postMaker, "gag/", previousGagDir, "genome.mrna.fasta", sep = ""),
                  " ", paste(postMaker, SppAbbrv, ".mrna_gag_edited.fasta", sep = ""), sep = ""))
cmd[1]
sapply(cmd, function(x) system(x))
```

Add the paths to the edited gff and fasta files to the metadataAssemblies table:
```{r}
metadataAssemblies$fixed_gag_gff <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, "_gag_edited.gff", sep = "")
metadataAssemblies$fixed_gag_fsa <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, "_gag_edited.fsa", sep = "")
metadataAssemblies$fixed_gag_tbl <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".tbl", sep = "")
metadataAssemblies$fixed_gag_mrna <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".mrna_gag_edited.fasta", sep = "")
metadataAssemblies$fixed_gag_proteins <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".proteins_gag_edited.fasta", sep = "")
```

Editing the Fustric.proteins_gag_edited.fasta that were generated with GAG:    
Copy proteins fasta to give new name:  
$ cp Fustric.proteins_gag_edited.fasta Fustric.proteins_gag_edited.noStops.fasta   

*** Over here: I moved the edits to remove the stop codons that were later in the script to this part - perhaps start
running the script over from here?   
Make changes to the proteins fasta to look like the original Maker output:  
$ sed -i 's/>protein|/>/g' Fustric.proteins_gag_edited.noStops.fasta   
$ sed -i 's/ ID=.*$//g' Fustric.proteins_gag_edited.noStops.fasta     
$ sed -i 's/*//g' Fustric.proteins_gag_edited.noStops.fasta  
Take a look:  
$ head -5 Fustric.proteins_gag_edited.noStops.fasta  

If all is good, copy it up to the postprocessing directory:  
$ mv Fustric.proteins_gag_edited.noStops.fasta Fustric.proteins_gag_edited.fasta   
  
Generate the ID mapping file:
```{r}
# Set Fustric Locus_tag prefix
locus_tag_prefixes <- "Fustric"

metadataAssemblies$locusTagPrefix <- locus_tag_prefixes

justify = 6
abrvGene = "G"
abrvTrans = "T"
cmd <- with(metadataAssemblies,
            paste("perl ", makerMapPath, " --prefix ", paste(locusTagPrefix, "_", sep = ""), 
                  " --justify ", justify, " --abrv_gene ", abrvGene, " --abrv_tran ", abrvTrans, 
                  " ",  fixed_gag_gff, 
                  " > ", paste(postMaker, SppAbbrv, ".map", sep = "")))
cmd[1]
sapply(cmd, function(x) system(x))
```

Remove the -RA from the mapping IDs: 
```{r}
cmd <- with(metadataAssemblies, paste("sed -i 's/-RA//g' ", paste(postMaker, SppAbbrv, ".map", sep = "")))
cmd[1]
sapply(cmd, function(x) system(x))
```

Complete the in-place renaming of the gff:  
```{r}
cmd <- with(metadataAssemblies, paste("cp ", fixed_gag_gff, 
                                      " ", paste(postMaker, SppAbbrv, ".Gag.fixedNames.gff", sep = ""), " ;",
                                      " ", makerMapGffPath, " ", paste(postMaker, SppAbbrv, ".map", sep = ""),
                                      " ", paste(postMaker, SppAbbrv, ".Gag.fixedNames.gff", sep = ""), sep = ""))
cmd[1]
sapply(cmd, function(x) system(x))
```

Complete the in-place renaming of the fasta files:  
```{r}
cmd <- with(metadataAssemblies, paste("cp ", fixed_gag_mrna, 
                                      " ", paste(postMaker, SppAbbrv, ".mrna.Gag.fixedNames.fasta", sep = ""), " ;",
                                      " ", makerMapFasta, " ", paste(postMaker, SppAbbrv, ".map", sep = ""),
                                      " ", paste(postMaker, SppAbbrv, ".mrna.Gag.fixedNames.fasta", sep = ""), sep = ""))
cmd[1]
sapply(cmd, function(x) system(x))
cmd2 <- with(metadataAssemblies, paste("cp ", fixed_gag_proteins, 
                                      " ", paste(postMaker, SppAbbrv, ".proteins.Gag.fixedNames.fasta", sep = ""), " ;",
                                      " ", makerMapFasta, " ", paste(postMaker, SppAbbrv, ".map", sep = ""),
                                      " ", paste(postMaker, SppAbbrv, ".proteins.Gag.fixedNames.fasta", sep = ""), sep = ""))
cmd2[1]
sapply(cmd2, function(x) system(x))

cmd3 <- with(metadataAssemblies, paste("cp ", paste(postMaker, "Fustric_rnd3.all.maker.trnascan.transcripts.fasta", sep = ""), 
                                      " ", paste(postMaker, SppAbbrv, ".trna.fixedNames.fasta", sep = ""), " ;",
                                      " ", makerMapFasta, " ", paste(postMaker, SppAbbrv, ".map", sep = ""),
                                      " ", paste(postMaker, SppAbbrv, ".trna.fixedNames.fasta", sep = ""), sep = ""))
cmd3[1]
sapply(cmd3, function(x) system(x))
```
Add the paths to the edited gff with the newly mapped names to the metadataAssemblies table:
```{r}
metadataAssemblies$gag_NewNames_gff <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".Gag.fixedNames.gff", sep = "")
metadataAssemblies$gag_NewNames_protein <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".proteins.Gag.fixedNames.fasta",
                                                 sep = "")
metadataAssemblies$gag_NewNames_mrna <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".mrna.Gag.fixedNames.fasta",
                                              sep = "")
```

Remove instances of ;Name=; in the gff:
```{r}
cmd <- with(metadataAssemblies, paste("sed -i 's/;Name=;/;/g' ", gag_NewNames_gff, sep = ""))
cmd[1]
sapply(cmd, function(x) system(x))
```

Fix mRNA headers:
```{r}
cmd <- with(metadataAssemblies, paste("sed -i 's/ ID=.*$//g' ", gag_NewNames_mrna, sep = ""))
cmd[1]
sapply(cmd, function(x) system(x))
```

Use GAG to generate the final NCBI tbl for tbl2asn:  
```{r}
cmd <- with(metadataAssemblies, paste("cd ", postMaker, " && ", gagPath, 
                                      " -f ", fixed_gag_fsa,
                                      " -g ", gag_NewNames_gff, 
                                      " -o gag_2_mappedNames_for_tbl2asn", sep = ""))
cmd[1]
sapply(cmd, function(x) system(x))
```

Note, protein name problems, gff Name= problems. Check how tbl2asn works at this point.  
May need to fix the protein names, and perform Name fixes as a last step instead of during earlier steps.  

$ cd gag_2_mappedNames_for_tbl2asn/  
$ mkdir tbl2asn  
$ cp genome.tbl tbl2asn/Fustric.tbl; cp genome.fasta tbl2asn/Fustric.fsa  
$ cp ../template.sbt tbl2asn/Fustric_template.sbt  
$ ~/prog/linux64.tbl2asn -j "[organism=Fusarium_tricinctum] [strain=FTRI INRA104]" -t Fustric_template.sbt -p. -M n -Z discrep -a r1k -V b -X E   

```{r}
metadataAssemblies$final_tbl <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".tbl", sep = "")
metadataAssemblies$final_gff <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".gff", sep = "")
metadataAssemblies$final_mrna <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".mrna.fasta", sep = "") 
metadataAssemblies$final_assembly <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".fsa", sep = "")
metadataAssemblies$final_proteins <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".proteins.fasta", sep = "")
metadataAssemblies$final_gbf <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".gbf", sep = "")
metadataAssemblies$final_trna <- paste(metadataAssemblies$postMaker, metadataAssemblies$SppAbbrv, ".trna.fasta", sep = "")
```
